### 2020-09-20 samplepaper.tex
Initial Overleaf Import

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage{booktabs}   %% For formal tables:
%% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{xcolor} %% For \color
\usepackage{todonotes} %% For \todo
\usepackage{mathtools} %% For \xRightarrow[sub]{sup}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd} %% For \llbracket etc.
\usepackage{semantic}
\usepackage[scaled]{beramono}
\usepackage[T1]{fontenc}
\usepackage{natbib} %% For more citation commames (\cite etc.)
\usepackage{listings} %% For typesetting code
\newcommand{\lstsetgrift}{
	\lstset{%
		language=Lisp,%
		basicstyle=\ttfamily\small,%
		keywordstyle=\ttfamily\bfseries\small,%
		columns=flexible,%
		aboveskip=\smallskipamount,%
		belowskip=\smallskipamount,%
		frame=single,%
		xleftmargin=2pt,%
		escapeinside={/+}{+/},%
		mathescape=true,%
		keywords=[1]{define,if,:,Int,Bool,-,=,open},
		literate=%
		{->}{{$\rightarrow$}}2%
		{=>}{{$\Rightarrow$}}2%
		{lambda}{{$\lambda$}}1%
		{→}{{$\rightarrow$ }}2%
		{Γ}{{$\Gamma$ }}1%
		{⊢}{{$\vdash$ }}1%
	}%
}%%
\lstsetgrift

% random stuff
\newcommand{\infr}[2]{\inference{#1}{#2}}
\newcommand{\highlight}[1]{{\color{red} #1}}
\newcommand{\sidecond}[1]{\text{if}\;#1}

% abbreviations in figures
\newcommand{\figreduceto}[3]{#1 & \longmapsto & #2 & #3 \\}
\newcommand{\figreducetoH}[3]{\highlight{#1} & \highlight{\longmapsto} & 
\highlight{#2} & \highlight{#3} \\}
\newcommand{\figequalto}[3]{#1 & = & #2 & #3 \\}
\newcommand{\figequaltoH}[3]{\highlight{#1} & \highlight{=} & \highlight{#2} & 
\highlight{#3} \\}
\newcommand{\stxinrule}[3]{\text{#2} & #1 & \in & #3\\}
\newcommand{\stxrule}[3]{\text{#2} & #1 & \Coloneqq & #3\\}
\newcommand{\stxrulecont}[1]{& & | & #1 \\}

% ref figures, sections, etc.
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\chapref}[1]{Section~\ref{#1}}

% names
\newcommand{\LUD}{\ensuremath{\text{Lazy\;UD}}}
\newcommand{\LD}{\ensuremath{\text{Lazy\;D}}}
\newcommand{\CastADT}{\ensuremath{\text{Cast ADT}}}
\newcommand{\RMachine}[1]{\ensuremath{\mathcal{R}(#1)}}
\newcommand{\LDMachine}{\RMachine{\BLD}}
\newcommand{\LUDMachine}{\RMachine{\BLUD}}
\newcommand{\RBMachine}[0]{\RMachine{B}}
\newcommand{\SMachine}[1]{\ensuremath{\mathcal{S}(#1)}}

% error monad
\newcommand{\just}[1]{\ensuremath{#1}}
\newcommand{\error}[1]{\ensuremath{\Uparrow#1}}

% types
\newcommand{\Tdyn}[0]{\ensuremath{\star}}

%% pre-types
\newcommand{\Pbool}[0]{\ensuremath{\mathtt{Bool}}}
\newcommand{\Pfunc}[2]{\ensuremath{#1 \shortrightarrow #2}}
\newcommand{\Pprod}[2]{\ensuremath{#1 \times #2}}

% ground types
\newcommand{\Gbool}[0]{\Pbool}
\newcommand{\Gfunc}[0]{\Pfunc{\Tdyn}{\Tdyn}}
\newcommand{\Gprod}[0]{\Pprod{\Tdyn}{\Tdyn}}

%% expressions
\newcommand{\econ}[1]{#1}
\newcommand{\evar}[1]{#1}
\newcommand{\elam}[3]{\lambda#1:#2.\;#3}
\newcommand{\eapp}[2]{#1\;#2}
\newcommand{\etrue}[0]{\mathtt{true}}
\newcommand{\efalse}[0]{\mathtt{false}}
\newcommand{\eif}[3]{\mathtt{if}\;#1\;#2\;#3}
\newcommand{\econs}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\ecar}[1]{\mathtt{car} \; #1}
\newcommand{\ecdr}[1]{\mathtt{cdr} \; #1}
%\newcommand{\ecast}[2]{#1 \langle #2 \rangle}
\newcommand{\ecast}[2]{\ensuremath{#1 : #2}}

% casts
\newcommand{\ccast}[3]{#1 \xRightarrow[]{\enskip #2 \enskip} #3}
% \Rightarrow^{#2}

% values
\newcommand{\vdyn}[2]{\mathtt{Dyn}_{#1}(#2)}
\newcommand{\vtrue}[0]{\mathtt{true}}
\newcommand{\vfalse}[0]{\mathtt{false}}
\newcommand{\vfunc}[4]{\langle\elam{#1}{#2}{#3},#4\rangle}
\newcommand{\vfuncc}[6]{\ecast{#1}{\ccast{\Pfunc{#2}{#3}}{#4}{\Pfunc{#5}{#6}}}}
\newcommand{\vcons}[2]{\econs{#1}{#2}}
\newcommand{\vconsc}[6]{\ecast{#1}{\ccast{\Pprod{#2}{#3}}{#4}{\Pprod{#5}{#6}}}}

% observables
\newcommand{\odyn}{\ensuremath{\mathtt{dynamic}}}
\newcommand{\otrue}[0]{\mathtt{true}}
\newcommand{\ofalse}[0]{\mathtt{false}}
\newcommand{\ofunc}[0]{\mathtt{function}}
\newcommand{\ocons}[0]{\mathtt{cons}}
\newcommand{\oerror}[1]{\error{#1}}

% blame strategies
\newcommand{\BLD}[0]{\ensuremath{\mathtt{LD}}}
\newcommand{\BLUD}[0]{\ensuremath{\mathtt{LUD}}}

% cast _r_esults
\newcommand{\rsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rfail}[1]{\mathtt{fail}\;#1}

% machine _s_tates
\newcommand{\sexpr}[3]{\ensuremath{\langle#1,#2,#3\rangle}}
\newcommand{\scont}[2]{\ensuremath{\langle#1,#2\rangle}}
\newcommand{\shalt}[1]{\ensuremath{\Downarrow #1}}

% continuations
\newcommand{\Knil}[0]{\Box}
\newcommand{\Kcons}[2]{[#1]#2}

% frames
\newcommand{\closure}[2]{\langle#1,#2\rangle}
\newcommand{\FappXO}[2]{\eapp{\square}{\closure{#1}{#2}}}
\newcommand{\FappOX}[1]{\eapp{#1}{\square}}
\newcommand{\Fif}[3]{\eif{\square}{\closure{#1}{#3}}{\closure{#2}{#3}}}
\newcommand{\FconsXO}[2]{\econs{\square}{\closure{#1}{#2}}}
\newcommand{\FconsOX}[1]{\econs{#1}{\square}}
\newcommand{\Fcar}[0]{\ecar{\square}}
\newcommand{\Fcdr}[0]{\ecdr{\square}}
\newcommand{\Fcast}[1]{\ecast{\square}{#1}}

% threesomes
\newcommand{\tscc}[3]{\ccast{#1}{#2}{#3}}
%\newcommand{\tsTp}[2]{({#1};{#2})}
\newcommand{\tsTp}[2]{{#2}^{#1}}
\newcommand{\tsTdyn}[0]{\Tdyn}
\newcommand{\tsPbool}[0]{\Pbool}
\newcommand{\tsPfunc}[2]{\Pfunc{#1}{#2}}
\newcommand{\tsPprod}[2]{\Pprod{#1}{#2}}
\newcommand{\tsPfail}[2]{\bot^{#1#2}}
\newcommand{\tsle}[0]{\epsilon}
%\newcommand{\tsll}[1]{\cnfproj{}{#1}}
\newcommand{\tsll}[1]{#1}

% coercions in normal form
\newcommand{\cnfproj}[2]{#1?^{#2}}
\newcommand{\cnfinj}[1]{#1!}
\newcommand{\cnfseq}[2]{#1;#2}
\newcommand{\cnffail}[1]{\bot^{#1}}
\newcommand{\cnffunc}[2]{\Pfunc{#1}{#2}}
\newcommand{\cnfprod}[2]{\Pprod{#1}{#2}}
\newcommand{\cnfid}[1]{\ensuremath{\mathtt{id}_{#1}}}

% hypercoercions
\newcommand{\hcci}[0]{\cnfid{\Tdyn}}
\newcommand{\hccc}[3]{\ensuremath{(#1;#2;#3)}}
\newcommand{\hche}[0]{\ensuremath{\epsilon}}
\newcommand{\hchp}[2]{\ensuremath{#1?^{#2}}}
\newcommand{\hcmbool}[0]{\Pbool}
\newcommand{\hcmfunc}[2]{\Pfunc{#1}{#2}}
\newcommand{\hcmprod}[2]{\Pprod{#1}{#2}}
\newcommand{\hcmfail}[1]{\ensuremath{\bot^l}}
\newcommand{\hcte}[0]{\ensuremath{\epsilon}}
\newcommand{\hcti}[1]{\ensuremath{#1!}}

% gaps
\newcommand{\gnone}{\ensuremath{\epsilon}}
\newcommand{\gsome}[3]{\ccast{#1}{#2}{#3}}

% operators
\newcommand{\mbind}[0]{\ensuremath{\;\gg=\;}}
\newcommand{\mcomp}[0]{\ensuremath{\;>=>\;}}

% functions
\newcommand{\meet}[2]{\ensuremath{#1 \sqcap #2}}
\newcommand{\denote}[1]{\ensuremath{\llbracket #1 \rrbracket}}
\newcommand{\proxy}[2]{\ensuremath{\mathit{proxy}(#1,#2)}}
\newcommand{\applyCast}[2]{\ensuremath{\mathit{applyCast}(#1,#2)}}
\newcommand{\project}[4]{\ensuremath{\mathit{project}(#1,#2,#3,#4)}}
\newcommand{\composegap}[3]{\ensuremath{#1 \fatsemi #2 \fatsemi #3}}
\newcommand{\mkhead}[2]{\ensuremath{\mathit{mkProj}(#1,#2)}}
\newcommand{\mktail}[2]{\ensuremath{\mathit{mkInj}(#1,#2)}}
\newcommand{\mkgap}[3]{\ensuremath{\mathit{mkGap}(#1,#2,#3)}}
\newcommand{\compose}[2]{\ensuremath{#1 \fatsemi #2}}
\newcommand{\translate}[1]{\ensuremath{\lceil#1\rceil}}
\newcommand{\castToDyn}[2]{\ensuremath{\mathit{castToDyn}(#1,#2)}}
\newcommand{\castFromDyn}[2]{\ensuremath{\mathit{castFromDyn}(#1,#2)}}
\newcommand{\dom}[1]{\ensuremath{\mathit{dom}(#1)}}
\newcommand{\cod}[1]{\ensuremath{\mathit{cod}(#1)}}
\newcommand{\car}[1]{\ensuremath{\mathit{car}(#1)}}
\newcommand{\cdr}[1]{\ensuremath{\mathit{cdr}(#1)}}
\newcommand{\id}[1]{\ensuremath{\mathit{id}(#1)}}
\newcommand{\idm}[1]{\ensuremath{\mathit{id}_m(#1)}}
\newcommand{\idc}[1]{\ensuremath{\mathit{id}_c(#1)}}
\newcommand{\mkCont}[1]{\ensuremath{\mathit{mkCont}(#1)}}
\newcommand{\extCont}[2]{\ensuremath{\mathit{extCont}(#1,#2)}}
\newcommand{\applyCont}[2]{\ensuremath{\langle\langle#1,#2\rangle\rangle}}

% other relations
\newcommand{\tailable}[2]{\ensuremath{\casttyping{!}{#1}{#2}}}
\newcommand{\consistent}[2]{\ensuremath{#1\sim#2}}
\newcommand{\shallowlyconsistent}[2]{\ensuremath{#1\smile#2}}
\newcommand{\notshallowlyconsistent}[2]{\ensuremath{#1\not\smile#2}}
\newcommand{\expressiontyping}[3]{\ensuremath{#1 \vdash #2 : #3}}
\newcommand{\matching}[2]{\ensuremath{#1 \triangleright #2}}
\newcommand{\valuetyping}[2]{\ensuremath{\vdash #1 : #2}}
\newcommand{\envtyping}[2]{\ensuremath{#1 \vdash #2}}
\newcommand{\casttyping}[3]{\ensuremath{\vdash #1 : #2 \Longrightarrow #3}}
\newcommand{\transitivelyreducetoM}[3]{\ensuremath{#2 \longmapsto_{#1}^{*} #3}}
\newcommand{\transitivelyreduceto}[2]{\ensuremath{#1 \longmapsto^{*} #2}}
\newcommand{\transitivelyreducetoD}[2]{
	\ensuremath{#1 \longmapsto_{\LDMachine}^{*} #2}}
\newcommand{\transitivelyreducetoUD}[2]{
	\ensuremath{#1 \longmapsto_{\LUDMachine}^{*} #2}}
\newcommand{\reduceto}[2]{\ensuremath{#1 \longmapsto #2}}
\newcommand{\reducetoR}[3]{\ensuremath{#2 \longmapsto_{\RMachine{#1}} #3}}
\newcommand{\reducetoD}[2]{\reducetoR{\BLD}{#1}{#2}}
\newcommand{\reducetoUD}[2]{\reducetoR{\BLUD}{#1}{#2}}
\newcommand{\reducetoS}[3]{\ensuremath{#2 \longmapsto_{\SMachine{#1}} #3}}
\newcommand{\evalto}[2]{\ensuremath{\mathit{eval}(#1)=#2}}
\newcommand{\withmachineevalto}[3]{\ensuremath{\mathit{eval}_{#1}(#2)=#3}}
\newcommand{\evaltoR}[3]{\ensuremath{\mathit{eval}_{\RMachine{#1}}(#2)=#3}}
\newcommand{\evaltoD}[2]{\evaltoR{\BLD}{#1}{#2}}
\newcommand{\evaltoUD}[2]{\evaltoR{\BLUD}{#1}{#2}}
\newcommand{\evaltoS}[3]{\ensuremath{\mathit{eval}_{\SMachine{#1}}(#2)=#3}}
\newcommand{\compatibletailhead}[2]{\ensuremath{#1\;\mathit{R}\;#2}}
\newcommand{\goodgap}[1]{\ensuremath{\mathit{good}(#1)}}
\newcommand{\compileto}[4]{\expressiontyping{#1}{#2\rightsquigarrow#3}{#4}}
\newcommand{\machineequiv}[2]{
	\ensuremath{
		\withmachineevalto{#1}{e}{o}	
		\text{ if and only if }
		\withmachineevalto{#2}{e}{o}
		\text{ for all }
		\expressiontyping{\emptyset}{e}{T}
		\text{ and }
		\valuetyping{o}{T}
	}}
\newcommand{\bisimE}[3]{\ensuremath{#1 \vdash #2 \approx #3}}
\newcommand{\bisimv}[3]{\valuetyping{#1 \approx #2}{#3}}
\newcommand{\bisimr}[3]{\valuetyping{#1 \approx #2}{#3}}
\newcommand{\bisimF}[4]{\casttyping{#1 \approx #2}{#3}{#4}}
\newcommand{\bisimc}[4]{\casttyping{#1 \approx #2}{#3}{#4}}
\newcommand{\bisimK}[4]{\casttyping{#1 \approx #2}{#3}{#4}}
\newcommand{\bisims}[3]{\valuetyping{\ensuremath{#1 \approx #2}}{#3}}

\begin{document}
%
\title{A Framework for Equivalence of Cast Calculi and a new Cast Representation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Kuang-Chen Lu\inst{1} \and
Jeremy G. Siek\inst{1} \and
Andre Kuhlenschmidt\inst{1}
}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Indiana University, Bloomington IN 47405, USA}
% \institute{Princeton University, Princeton NJ 08544, USA \and
% Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
% \email{lncs@springer.com}\\
% \url{http://www.springer.com/gp/computer-science/lncs} \and
% ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
% \email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Gradual typing aims at combining dynamic typing and static typing in one 
language to provide the advantages of both typing disciplines. Implementing 
gradual typing with reasonable overhead, however, is not trivial. Casts might 
accumulate in higher-order values and continuations, thus change the overall 
space complexity. Several space-efficient cast representations have been 
proposed, including coercions in normal forms, threesomes, and supercoercions. 
Drawing an interface between a cast representation and the remaining parts of 
language implementations would likely help to understand the relation 
between these cast representations and potentially new ones. We present such an 
interface, Cast Abstract Data Type. The Cast ADT brings another 
benefit. Space-efficient cast representations shall be proved not changing the 
overall behavior of language implementations except saving space. Existing 
proofs typically go through a whole bisimulation proof between the standard and 
inefficient semantics and a space-efficient one. If a space-efficient semantics 
uses the Cast ADT, its behavior would depend solely on the properties of the 
interface functions. Thus a cast representation is correct if its interface 
functions work correctly. We present an abstract machine \SMachine{C} that uses 
Cast ADT to work with casts, a parameterized bisimulation proof, and two 
subsets of Cast ADT that specify the Lazy D and the Lazy UD blame strategies. We
also introduce a new cast representation, hypercoercions, that is better than 
the known cast representations in some aspects.

% The abstract should briefly summarize the contents of the paper in
% 15--250 words.

\keywords{Gradual Typing \and Blame \and Coercion.}
\end{abstract}
%
%
%

\section{Introduction}
Around 2006, several groups of researchers proposed ways to integrate
dynamic typing and static typing, notably gradual typing
\citep{siek2006gradual}, hybrid typing \citep{knowles2010hybrid},
migratory typing \citep{tobin2006interlanguage}, and multi-language
interoperability \citep{Gray:2005ij,Matthews:2007zr}.

\todo[inline]{Cite "Contracts for higher-order functions" for the proxy-based approach?}

Gradual typing \citep{siek2006gradual,siek2015refined} includes both static typing 
and dynamic typing in the same language, enables smooth transition between them, while
enforcing type soundness.
To allow the smooth transition, gradually typed languages are featuerd by a special type
typically written as \Tdyn\ or \lstinline|Dyn| or \lstinline|Any|, which enables 
imprecise type annotations. A gradual type checker would allow programmers to cast 
implicitly between types that have no conflicting type information.
Researchers usually define the semantics of gradually typed languages by
translation to an intermediate language with explicit casts, such as the blame
calculus \citep{wadler2009well} and other cast calculi \citep{siek2009exploring}.
At runtime, casts store or check type information to enforce type soundness.
% To allow the smooth transition, every gradually typed language
% has a special type typically written as \Tdyn\ or \lstinline|Dyn| or \lstinline|Any|,
% which enables imprecise type annotations.
% Programmers can use an expression of type \texttt{T} in any context that \texttt{T} 
% is consistent to.
% Two types are consistent if they have no conflicting type information.
% For instance, \lstinline|(->  Bool Int)| is consistent with \lstinline|(->  $\star$ Int)| but
% not consistent with \lstinline|(->  Int Int)|.
% When the type of an expression is consistent but not equal to the type of its 
% context, there would be a runtime computation (i.e. a cast) that store or check 
% type information to enforce type soundness.

Unfortunately, straightforward implementations of casts impose significant 
runtime overheads that can change the asymptotic space complexity of a program 
\citep{herman2010space}.
For instance, the following module define functions 
that check the parity of natural numbers. The terms are ordinary, while the 
type annotations are not -- the output type of \lstinline|even?| is 
imprecise.
\begin{lstlisting}
(define (even? [n : Int]) : $\star$
  (if (= n 0) #t (odd? (- n 1))))
(define (odd? [n : Int]) : Bool
  (if (= n 0) #f (even? (- n 1))))
\end{lstlisting}
These functions look tail-recursive, so it is natural to expect that they 
consumes constant space. This is not the case, however, in a straightforward 
implementation because there is a cast around each of the 
``tail call'', one from \lstinline|Bool| to 
\lstinline|$\star$|, the other goes the other way.
A straightforward implementation would cast the output of the recursive 
call \textit{before} returning it, which effectively breaks the tail recursion 
and thus changes the space complexity.

Continuation is not the only place where casts might accumulate. 
Consider a variant of the above program written in continuation-passing style. 
Here, both functions returns \lstinline|Bool| so there is not pending cast in 
continuations. The continuation argument \lstinline|k|, however, is wrapped in 
a cast in each recursive call because the type of \lstinline|k| is less 
precise in \lstinline|even?|.
\begin{lstlisting}
(define (even? [n : Int] [k : (-> $\star$ Bool)]) : Bool
  (if (= n 0) (k #t) (odd? (- n 1) k)))
(define (odd? [n : Int] [k : (-> Bool Bool)]) : Bool
  (if (= n 0) (k #f) (even? (- n 1) k)))
\end{lstlisting}
Our examples programs are inspired by the examples in \citep{herman2010space}.
We direct interested readers to their work for more examples and detailed explanation.

A solution to this space-efficiency problem is representing cast information 
more concisely. There are several known space-efficient cast representations
\citep{Siek:2015:BCT:2737924.2737968,Siek:2010:TWB:1706299.1706342,Garcia:2013:CTB:2500365.2500603,kuhlenschmidt2018efficient,siek2012interpretations,garcia2014deriving}.
The current state of the art includes
\begin{itemize}
	\item threesomes 
	\citep{Siek:2010:TWB:1706299.1706342,Garcia:2013:CTB:2500365.2500603},
	\item supercoercions \citep{Garcia:2013:CTB:2500365.2500603}, and
	\item coercions in normal form
	\citep{siek2012interpretations,Siek:2015:BCT:2737924.2737968}.
\end{itemize}

These cast representations are represented in their own dynamic semantics of a gradually typed language,
which make it harder to focus on key differences between these representations. 
Surprisingly, no one has published a dynamic semantics parameterized over cast representations.
We present in Section~\ref{sec:framework} first of such generic dynamic semantics in the form 
of a CEK abstract machine and formulate the interface between the machine and cast representations
as an abstract data type, namely Cast ADT.

Of course, an alternative cast representation must be proved
correct. That is, a language implementation should give the same output no 
matter whether it is using a cast representation to reduce space 
consumption. Our generic CEK machine comes with generic theorems (Section~\ref{sec:framework}),
with which researchers can prove a cast representation is correct by showing its corresponding
implementation of Cast ADT satisfies a bunch of properties. We believe this framework can reduce the burden
 of proving this kind of theorems and help people focusing on the cast representations.
For illustration, we apply the framework to Lazy UD coercions in normal form
and Lazy UD threesomes (Chapter~\ref{sec:framework-examples}). 
We believe the framework is also applicable to other cast representations.

\todo[inline]{Actuallythink twosome's composition is structurally recursive but threesomes' is not}
The published cast representations have various strengths and weaknesses.
The major insight of these systems is a compose operator for casts which compress cast information to
reduce space consumption.
Threesomes and supercoercions are good for mechanized
metatheory because their compose operators are structurally recursive,
making them easy to define in proof assistants such as Agda. In
contrast, the coercions in normal form have compose operators that are
not structurally recursive, which makes it more difficult to define in
Agda, requiring what amounts to an explicit proof of termination.
%
On the other hand, coercions in normal form are easier to understand
than threesomes (with a strange labeled bottom type), and
supercoercions (10 different kinds).

This paper presents a new cast representation 
(Section~\ref{chap:hypercoercions}), named
\emph{hypercoercions}, that is good for both mechanized metatheory and
good for implementation. The straightforward Agda translation of composition 
operator for hypercoercions is able to pass the termination check and 
hypercoercions are suggestive of a bit-level representation that minimizes the 
need for pointers and fits all first-order casts into 64 bits.
%
We present two flavors of hypercoercions to support the two blame
tracking strategies from the literature: \LD\ and
\LUD~\citep{siek2009exploring}. With the \LD\ strategy, only 
\textbf{d}own 
casts (casts from \Tdyn\ to some other types) are subject to blame. 
With the 
\LUD\
strategy, however, some \textbf{u}p casts are also blamable.
We are interested in the D blame
tracking strategy because it comes with a more straightforward notion of
safe cast compared to \LUD~\citep{siek2009exploring}, which is why \LD\ 
was
chosen in the Grift compiler~\citep{kuhlenschmidt2018efficient}. We are also 
interested in \LUD\ because it plays a
prominent role in the gradual typing literature
\citep{wadler2009well}.  The \LUD\ hypercoercions were inspired by the
supercoercions of \citep{Garcia:2013:CTB:2500365.2500603} (hence the name) and
the \LD\ hypercoercions were inspired by the normal forms of
\citep{siek2012interpretations}.
The semantics of casts can be lazy or
eager~\citep{siek2009exploring}. In this paper, we focus on lazy cast
strategies because we suspect that they are more efficient than eager
strategies and because \citet{new2019gradual} show that the eager
strategies are incompatible with $\eta$-equivalence of functions.

To summarize, the primary contributions of this paper are:
\begin{itemize}
    \item Cast ADT, an abstract data type for cast representations
    \item a CEK abstract machine parameterized over Cast ADT
	\item a framework in Agda for proving the correctness of \LD\ or \LUD\ 
	cast representations when plugging into the CEK machine
	\item a formal proof that \LUD\ threesomes are correct
	\item a formal proof that \LUD\ coercions in normal form are correct
	\item hypercoercions, a new cast representation, which
	has a more compact memory representation in comparison to coercions in 
	normal form.
	\item a formal proof that both \LUD\ hypercoercions are correct
	\item a formal proof that both \LD\ hypercoercions are correct
\end{itemize}

\section{Related Works}
\label{chap:related-works}

\todo[inline]{how to make citet work?}

\subsubsection{Coercions in normal form}

\citet{henglein1994dynamic} introduces coercions to study compilation from 
dynamic languages to static ones. \citet{herman2010space} proposes using 
coercions in normal forms to represent casts in gradual typing and proves space 
efficiency. 
However, they didn't show the coercion-based semantics is equivalent to the 
standard (and type-based) one. \citet{Siek:2015:BCT:2737924.2737968} improves 
the Lazy UD flavor of coercions in normal forms and proves that Lazy UD 
coercions in normal form is correct, using two bisimulation between three 
reduction semantics. The first semantics is similar to \LUDMachine, which 
bisimulates a 
reduction semantics using coercions (not necessarily in normal form). The last 
semantics uses coercions in normal form and bisimulates the coercions-based 
semantics.

\subsubsection{Threesomes}

\citet{Siek:2010:TWB:1706299.1706342} introduces Lazy UD threesomes and prove 
its correctness, using a bisimulation between two reduction semantics: one is 
similar to \LUDMachine; the other uses Lazy UD threesomes and is similar to the 
machine in Section~\ref{sec:framework}. 
\citet{Garcia:2013:CTB:2500365.2500603} introduces several other flavors of 
threesomes, including a Lazy D one. Their threesomes are derived from 
corresponding flavors of coercions in normal form, using correctness-preserving 
transformation. So the correctness of their threesomes relies on the 
corrections of the coercions-in-normal-form-based semantics defined in 
\citet{herman2010space}, which have not been proved correct.

\subsubsection{Parameterized cast calculi}

\citet{garcia2014deriving} presents a dynamic semantics for the Eager D blame 
strategy, using coercions. They claims that "the artefacts for coercions can be
replaced by other artefacts implementing different dynamic semantics".

\section{Background} \label{sec:background}

In this section, we first review lazy cast calculi (Section~\ref{sec:cast-calculus}), where
we present an abstract machine that is employed in our framework \ref{sec:framework}. Then
we review coercions and their normal forms (Section~\ref{sec:coercions-in-normal-form}), which
motivates and inspires the design of hypercoercions (Section~\ref{sec:hypercoercions}).

\subsection{Cast Calculi} \label{sec:cast-calculi}

\begin{figure}
Syntax
\[
\begin{array}{lrcl}
\text{Blame Labels} & l \\
\text{Variables} & x,y,z \\
\stxrule{S,T}{Types}{
	\Tdyn \mid{}
	P
}
\stxrule{P,Q}{Pre-types}{
	\Pbool \mid
	\Pfunc{T}{T} \mid
	\Pprod{T}{T}
}
\stxrule{G,H}{Ground types}{
	\Gbool  \mid
	\Gfunc  \mid
	\Gprod
}
\stxrule{e}{Expressions}{
	\evar{x}        \mid
	\elam{x}{S}{e}  \mid
	\eapp{e}{e}     \mid
	\econs{e}{e}  \mid
	\ecar{e}      \mid
	\ecdr{e}
}
\stxrulecont{
	\ecast{e}{\ccast{S}{l}{T}}  \mid
	\etrue         \mid
	\efalse        \mid
	\eif{e}{e}{e}
}
\end{array}
\]

	\fbox{\consistent{T}{T}} Consistency
	\begin{gather*}
	\infr{}{\consistent{\Tdyn}{\Tdyn}} \quad
	\infr{}{\consistent{\Tdyn}{P}    } \quad
	\infr{}{\consistent{P}{\Tdyn}    }
	\\[1ex]
	\infr{}{
		\consistent{\Pbool}{\Pbool}
	} \quad
	\infr{
		\consistent{S_1}{S_2} &
		\consistent{T_1}{T_2}
	}{
		\consistent{\Pfunc{S_1}{T_1}}{\Pfunc{S_2}{T_2}}
	} \quad
	\infr{
		\consistent{S_1}{S_2} &
		\consistent{T_1}{T_2}
	}{
		\consistent{\Pprod{S_1}{T_1}}{\Pprod{S_2}{T_2}}
	}
	\end{gather*}

	\fbox{\shallowlyconsistent{T}{T}} Shallow consistency
	\begin{gather*}
	\inference{}{
		\shallowlyconsistent{\Tdyn}{\Tdyn}
	} \quad
	\inference{}{
		\shallowlyconsistent{\Tdyn}{P}
	} \quad
	\inference{}{
		\shallowlyconsistent{P}{\Tdyn}
	} \\
	\inference{}{
		\shallowlyconsistent{\Pbool}{\Pbool}
	} \quad
	\inference{}{
		\shallowlyconsistent{\Pfunc{S_1}{T_1}}{\Pfunc{S_2}{T_2}}
	} \quad
	\inference{}{
		\shallowlyconsistent{\Pprod{S_1}{T_1}}{\Pprod{S_2}{T_2}}
	}
	\end{gather*}

\fbox{$\expressiontyping{\Gamma}{e}{T}$} Expression Typing
\begin{gather*}
\dots \quad
\infr{
	\expressiontyping{\Gamma}{e}{S}  &
	\consistent{S}{T}
}{
	\expressiontyping{\Gamma}{(\ecast{e}{\ccast{S}{l}{T}})}{T}
}
\end{gather*}

\caption{Syntax and static semantics of the Cast Calculus.}
\label{fig:statics}
\end{figure}

\paragraph{Syntax and Static Semantics}

The syntax and static semantics are the same for the \LD\ and
\LUD\ cast calculi. They are reviewed in \figref{fig:statics}.
As usual, the important features are the cast expressions,
$\ecast{e}{\ccast{S}{l}{T}}$, which are responsible for runtime type
checking. Polarities of blame labels is not treated but are straightforward to 
incorporate. The syntax and static semantics is the same as that of 
\citep{siek2009exploring} except for a few minor exceptions:

\begin{itemize}
\item We add product type.
\item We consider only one base type, Boolean, for simplicity.
\item We separate types into those with a type constructor at the top,
  the \emph{pre-types}, versus the dynamic type \Tdyn.
\end{itemize}
As usual, for every cast $\ecast{e}{\ccast{S}{l}{T}}$, the source type $S$ and 
the target $T$ must be consistent.

\paragraph{Dynamic Semantics}

The dynamic semantics of a cast calculus is typically defined with a
reduction semantics. Here we use a CEK machine
\citep{felleisen1986control} instead. One reason is that environments are 
easier 
to deal with than substitutions in formal proof. Also, a CEK machine is more
convenient to use for the space-efficient semantics in
\chapref{sec:framework} (recall that a major point of space-efficiency 
is about compressing continuations and that continuations are part of the CEK 
machine states). So using a CEK machine for the cast 
calculi in this section makes it easier to prove the equivalent of these 
machines. Of course, one should prove that the abstract machine
presented here is equivalent to the standard reduction semantics for
cast calculi, but we have not yet done so.

\begin{figure}
\[
\begin{array}{lrcl}
\stxrule{v,u}{Values}{
  	\vdyn{I}{v}  \mid
    \vtrue       \mid
    \vfalse                     
}
\stxrulecont{
    \vfunc{x}{S}{e}{E}  \mid
	\vfuncc{v}{S}{T}{l}{S}{T}
}
\stxrulecont{
    \vcons{v}{v}               \mid
	\vconsc{v}{S}{T}{l}{S}{T}
}
\stxrule{I,J}{Injectable types}{
	\text{a subset of $P$ defined by the blame strategy $B$}
}
\stxrule{E}{Environments}{
	\text{a partial function}\;\{\langle x , v \rangle , \dots \}
}
\stxrule{F}{Frames}{
	\FappXO{e}{E}  \mid
	\FappOX{v}               \mid
	\Fcast{\ccast{S}{l}{T}}                \mid
	\Fcar                    
}
\stxrulecont{
	\Fcdr           \mid
	\FconsXO{e}{E}  \mid
	\FconsOX{v}
}
\stxrule{K}{Continuations}{
	\Knil         \mid
	\Kcons{F}{K}
}
\stxrule{s}{States}{
	\sexpr{e}{E}{K}  \mid
	\scont{v}{K}     \mid
	\shalt{v}        \mid
	\error{l}
}
\stxrule{o}{Observables}{
    \odyn    \mid
    \otrue   \mid
    \ofalse  \mid
    \ofunc   \mid
    \ocons   \mid
    \error{l}
}
\stxrule{r}{Cast results}{
    v \mid
    \error{l}
}
\end{array}
\]

\fbox{\valuetyping{v}{T}} Value typing
\begin{gather*}
\dots \qquad
\inference{
	\valuetyping{v}{I}
}{
	\valuetyping{\vdyn{I}{v}}{\Tdyn}
} \\[1ex]
\inference{
	\valuetyping{v}{\Pfunc{S_1}{T_1}}
}{
	\valuetyping{(\vfuncc{v}{S_1}{T_1}{l}{S_2}{T_2})}{\Pfunc{S_2}{T_2}}
} \quad
\inference{
	\valuetyping{v}{\Pprod{S_1}{T_1}}
}{
	\valuetyping{(\vconsc{v}{S_1}{T_1}{l}{S_2}{T_2})}{\Pprod{S_2}{T_2}}
}
\end{gather*}

\fbox{$\proxy{v}{\ccast{S}{l}{T}}=v$}
\[
\begin{array}{rclr}
\figequalto{\proxy{v}{\ccast{\Pbool}{l}{\Pbool}}}{v}{}
\figequalto{\proxy{v}{\ccast{\Pfunc{S_1}{T_1}}{l}{\Pfunc{S_2}{T_2}}}}{
	\vfuncc{v}{S_1}{T_1}{l}{S_2}{T_2}
}{}
\figequalto{\proxy{v}{\ccast{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}{
	\vconsc{v}{S_1}{T_1}{l}{S_2}{T_2}
}{}
\end{array}
\]

\caption{Syntax of a CEK machine \RMachine{B} and auxiliary definitions.}
\label{fig:R-machine-syntax}
\end{figure}

Fig.~\ref{fig:R-machine-dynamics} defines the transition relation of the CEK
machine (\RMachine{B}) and Fig.~\ref{fig:R-machine-syntax} gives a grammar for machine 
states $s$, including a definition of values $v$ and value typing.
%
This machine is parameterized over blame strategy ($B$), which ranges over
Lazy D (Fig.~\ref{fig:blame-strategy-LazyD}) and Lazy UD 
(Fig.~\ref{fig:blame-strategy-LazyD}).
%
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machines involve two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an evaluating
$\sexpr{e}{E}{K}$ or returning $\scont{v}{K}$ configuration, 
respectively. Additionally, there are two halting configurations: \shalt{v} 
halts with a value, while \error{l} halts with a blame label. 
Observables include all value constructors and blame labels, as in 
\citep{siek2012interpretations}. 
The function converting values to observables ($observe(v) = o$) is defined in 
the obvious way.

Let $v$ range over values. The 
$\vdyn{I}{v}$ stands for values casted to the dynamic type, as in 
\citep{wadler2009well}. The inner value must be of an 
injectable type. The definition of injectable types depends on blame 
strategies:
%
for the \LD\ strategy, every pre-type is injectable 
(Fig.~\ref{fig:blame-strategy-LazyD});
%
for the \LUD\ strategy, only ground types are injectable 
(Fig.~\ref{fig:blame-strategy-LazyUD}).
%
Other kinds of values are either Boolean constants, functions, casted 
functions, pairs, or casted pairs.
% In earlier version of this work, 
% we were tempted to use a generic constructor for all casted value, and use 
% value typing rules to restrict the cast such that its source and target are 
% shallowly consistent pre-types. This approach unfortunately considers 
% $\ecast{\etrue}{\ccast{\Pbool}{l}{\Pbool}}$ as a value.

The syntax \error{l} occurs several times in the syntax definition, we shall 
view those nonterminals that have \error{l} case as error monads. We use 
the symbol \mbind as the monad bind operator, which is defined as follows:
\[
\begin{array}{rcll}
(\error{l}) \mbind f & = & \error{l} \\
          a \mbind f & = & f(a)      & \text{otherwise}
\end{array}
\]

\begin{figure}
	\fbox{\reducetoR{B}{s}{s}} Transition
	\[
	\begin{array}{rclr}
	\figreduceto{\sexpr{\evar{x}}{E}{K}}{\scont{E(x)}{K}}{}
	\figreduceto{\sexpr{\elam{x}{S}{e}}{E}{K}}{\scont{\vfunc{x}{S}{e}{E}}{K}}{}
	\figreduceto{\sexpr{\eapp{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FappXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\etrue}{E}{K}}{\scont{\vtrue}{K}}{}
	\figreduceto{\sexpr{\efalse}{E}{K}}{\scont{\vfalse}{K}}{}
	\figreduceto{\sexpr{\eif{e_1}{e_2}{e_3}}{E}{K}}{
		\scont{e_1}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\econs{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FconsXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\ecar{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Fcar}{K}}}{}
	\figreduceto{\sexpr{\ecdr{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Fcdr}{K}}}{}
	\figreducetoH{
		\sexpr{\ecast{e}{\ccast{S}{l}{T}}}{E}{K}
	}{
		\sexpr{e}{E}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}}{
		\applyCast{v}{\ccast{S}{l}{T}} \; \mbind \; \lambda u. \scont{u}{K}
	}{}
	\figreduceto{\scont{v_1}{\Kcons{\FappXO{e_2}{E}}{K}}}{\sexpr{e_2}{E}{
			\Kcons{\FappOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v}{\Kcons{\FappOX{\vfunc{x}{S}{e}{E}}}{K}}}{
		\sexpr{e}{E[x:=v]}{K}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\FappOX{(\vfuncc{u}{S_1}{T_1}{l}{S_2}{T_2})}}{K}}}{
		\scont{v}{
			\Kcons{\Fcast{\ccast{S_2}{l}{S_1}}}{
				\Kcons{\FappOX{u}}{
					\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}}
	}{}
	\figreduceto{\scont{\vtrue}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_2}{E}{K}}{}
	\figreduceto{\scont{\vfalse}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_3}{E}{K}}{}
	\figreduceto{\scont{v_1}{\Kcons{\FconsXO{e_2}{E}}{K}}}{
		\sexpr{e_2}{E}{\Kcons{\FconsOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v_2}{\Kcons{\FconsOX{v_1}}{K}}}{
		\scont{\vcons{v_1}{v_2}}{K}
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Fcar}{K}}}{\scont{v_1}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Fcar}{K}}}{
		\scont{u}{
			\Kcons{\Fcar}{
				\Kcons{\Fcast{\ccast{S_1}{l}{S_2}}}{K}}}	
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Fcdr}{K}}}{\scont{v_2}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Fcdr}{K}}}{
		\scont{u}{
			\Kcons{\Fcdr}{
				\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}	
	}{}
	\figreduceto{\scont{v}{\Knil}}{\shalt{v}}{}
	\end{array}
	\]

	\fbox{\evaltoR{B}{e}{o}} Evaluation
	\[
	\inference{
		\transitivelyreduceto{\sexpr{e}{\emptyset}{\Knil}}{\shalt{v}}
	}{
		\evalto{e}{observe(v)}
	} \quad
	\infr{
		\transitivelyreduceto{\sexpr{e}{\emptyset}{\Knil}}{\error{l}}
	}{
		\evalto{e}{\error{l}}}
	\]
	\caption{Dynamic semantics of the cast calculi as a CEK
		machine \RMachine{B}. The transitions that involve casts are highlighted
		in red.}
	\label{fig:R-machine-dynamics}
\end{figure}

The transition relation \reduceto{s}{s} invokes the cast application function, 
noted as $\applyCast{v}{\ccast{S}{l}{T}}$ to allow for the differences 
between blame strategies (Fig.~\ref{fig:blame-strategy-LazyD} and 
Fig.~\ref{fig:blame-strategy-LazyUD}).
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%
To cast a value, the machine invokes the cast application function on the 
value. The cast result is either a value or an error (\error{l}). If the cast 
succeeds, the machine returns the value to the next continuation.  If the cast 
fails, the machine halts with the blame label.
%
To apply a casted function, the machine first casts the argument $v$, then
applies the inner function $u$ to the casted argument, and finally
casts the return value.
%
To take out the first part of a casted pair, the machine takes out the 
first part of the inner pair $u$ and cast it.
%
Taking out the second part is similar.

The reflexive transitive closure of reduction (\transitivelyreduceto{s}{s}) 
and evaluation (${\evalto{e}{o}}$) are standard~\citep{felleisen03:_pllc}.

\begin{figure}
\[
\begin{array}{lrcl}
\stxrule{I,J}{Injectable types}{P}
\end{array}
\]

\fbox{$\mathit{applyCast}_{\BLD}(v,\ccast{S}{l}{T}) = r$}
\[
\begin{array}{rclr}
\figequalto{
	applyCast(v,\ccast{\Tdyn}{l}{\Tdyn})
}{
	\just{v}
}{}
\figequalto{
applyCast(v,\ccast{P}{l}{\Tdyn})
}{
	\just{\vdyn{P}{v}}
}{}
\figequalto{
	applyCast(\vdyn{P}{u},\ccast{\Tdyn}{l}{Q})
}{
	applyCast(u,\ccast{P}{l}{Q})
}{}
\figequalto{
	applyCast(v,\ccast{P}{l}{Q})
}{
	\just{\proxy{v}{\ccast{P}{l}{Q}}}
}{\sidecond{\shallowlyconsistent{P}{Q}}}
\figequalto{
	applyCast(v,\ccast{P}{l}{Q})
}{
	\error{l}
}{\sidecond{\notshallowlyconsistent{P}{Q}}}
\end{array}
\]
\caption{The \LD\ blame strategy (\BLD)}
\label{fig:blame-strategy-LazyD}
\end{figure}

\begin{definition}[\LD\ CEK Machine, \LDMachine]
The \LD\ CEK machine, written \LDMachine, is the CEK machine of
\figref{fig:R-machine-dynamics} using the \BLD\ blame strategy defined in 
\figref{fig:blame-strategy-LazyD}.  We write the transition relations of
this machine as \reducetoD{s}{s} and \transitivelyreducetoD{s}{s}
and the evaluator as \evaltoD{e}{o}.
\end{definition}

We conjecture that \LDMachine\ agrees with the \LD\ cast calculus
of \citep{siek2009exploring}.

\begin{proposition}[\LDMachine\ is deterministic]
	If \reducetoD{s}{s_1} and \reducetoD{s}{s_2} then $s_1 = s_2$.
\end{proposition}

Next, we define the CEK Machine for \LUD. The only differences with
respect to \LD\ are in the injectable types and in the definition of the 
${\mathit{applyCast}(v,\ccast{S}{l}{T})}$
function, in which a cast whose source or target is the dynamic type
\Tdyn\ is always split into two casts that go through a ground type, that is, a 
type in which all sub-components are the dynamic type, such as 
\Pfunc{\Tdyn}{\Tdyn}.
\begin{definition}[\LUD\ CEK Machine]
	The \LUD\ CEK machine, written \LUDMachine\, is the CEK machine of
	\figref{fig:R-machine-dynamics} using the \BLUD\ blame strategy defined in 
	\figref{fig:blame-strategy-LazyUD}.  We write the transition relations of
	this machine as \reducetoUD{s}{s} and \transitivelyreducetoUD{s}{s}
	and the evaluator as \evaltoUD{e}{o}.
\end{definition}

\begin{figure}
\[
\begin{array}{lrcl}
\stxrule{I,J}{Injectable types}{G}
\end{array}
\]

	\fbox{$\mathit{applyCast}_{\mathtt{\BLUD}}(v,\ccast{S}{l}{T}) = r$}
	\[
	\begin{array}{rclr}
	\mathit{applyCast}(v, \ccast{\Tdyn}{l}{\Tdyn}) &=& \just{v} \\
	\mathit{applyCast}(v, \ccast{G}{l}{\Tdyn}) &=&
	\just{\vdyn{G}{v}} \\
	\mathit{applyCast}(v, \ccast{P}{l}{\Tdyn}) &=&
	\just{\vdyn{G}{\proxy{v}{\ccast{P}{l}{G}}}}
	& \text{if } G \sim P, G \neq P \\  
	\mathit{applyCast}(\vdyn{G}{u}, \ccast{\Tdyn}{l}{H}) &=&          
	\project{u}{G}{l}{H}\\
	\mathit{applyCast}(\vdyn{G}{u}, \ccast{\Tdyn}{l}{Q}) &=&
	\project{u}{G}{l}{H} \mbind \lambda v. \proxy{v}{\ccast{H}{l}{Q}}
	& \text{if } H \sim Q, H \neq Q \\
	\mathit{applyCast}(v, \ccast{P}{l}{Q}) &=&
	\just{\proxy{v}{\ccast{P}{l}{Q}}}
	& \text{if } P \smile Q \\
	\mathit{applyCast}(v, \ccast{P}{l}{Q}) &=&
	\error{l}
	& \text{if } P \not\smile Q
	\end{array}
	\]

	\fbox{$\project{v}{G}{l}{H}=r$}
	\[
	\begin{array}{rclr}
	\figequalto{\project{v}{G}{l}{G}}{v}{}
	\figequalto{\project{v}{G}{l}{H}}{\error{l}}{\sidecond{G \neq H}}
	\end{array}
	\]

\caption{The \LUD\ blame strategy (\BLUD)}
	\label{fig:blame-strategy-LazyUD}
\end{figure}

We conjecture that \LUDMachine\ agrees with the \LUD\ cast
calculus of \citep{siek2009exploring}.

\begin{proposition}[\LUDMachine\ is deterministic]
	If $\reducetoUD{s}{s_1}$ and $\reducetoUD{s}{s_2}$ 
	then $s_1 = s_2$.
\end{proposition}

\subsection{Coercions and Normal Forms} 
\label{sec:coercion-calculus}

In this section, we review the
coercions~\citep{henglein1994dynamic,herman2010space} and the normal form of
\citet{siek2012interpretations} to motivate the design of hypercoercions.  We
omit sum types and product types in this section because
\citet{siek2012interpretations} did not discuss them. We assume a basic
familiarity with coercions, and suggest that readers unfamiliar with coercions
to familiarize themselves with \citet{siek2012interpretations} and
\citet{Siek:2015:BCT:2737924.2737968}.

\begin{figure}
	\[
	\begin{array}{lrcl}
	\stxrule{I,J}{Injectable types (\LD)}{
		\Pfunc{T}{T} \mid \Pbool}
	\stxrule{I,J}{Injectable types (\LUD)}{
		\Pfunc{\Tdyn}{\Tdyn} \mid \Pbool
	}
	\stxrule{c,d}{Coercions}{
		\cnfid{T}       \mid
		\cnffunc{c}{d}  \mid
		\cnfprod{c}{d}  \mid
		\cnffail{l}     \mid
		\cnfinj{I}      \mid
		\cnfproj{I}{l}  \mid
		\cnfseq{c}{d}
	}
	\stxrule{s,t}{Coercions in normal form}{
		\cnfid{\Tdyn}       \mid
		\cnfseq{\cnfproj{I}{l}}{i} \mid
		i
	}
	\stxrule{i}{Final Coercions}{
		\cnffail{l} \mid
		\cnfseq{g}{\cnfinj{I}} \mid
		g
	}
	\stxrule{g}{Middle Coercions}{
		\cnfid{\Pbool} \mid
		\cnffunc{s}{t} \mid
		\cnfprod{s}{t}
	}
	\end{array}
	\]
	\caption{Syntax of coercions and normal forms {\`a} la 
		\citep{siek2012interpretations}.}
	\label{fig:normal-coercion}
\end{figure}

\figref{fig:normal-coercion} reviews the grammar for coercions,
written $c$.
%
To review, an injection $\cnfinj{I}$ takes a value from an injectable type
$I$ to type $\Tdyn$. An injectable type is simply a type that can be
cast directly to and from $\Tdyn$. The
definition of injectable types depends on the blame strategy, as in 
Chapter~\ref{chap:background}.
%
A projection $I?^l$ takes a value from type $\Tdyn$ to type $I$, or
halts the program and blames $l$ if the value is inconsistent with 
the target type.
The coercion \cnfid{T} is the identity, $\bot^{l}$ is the coercion that
always fails and blames $l$, and $(\cnfseq{c}{d})$ applies $c$ and
then $d$ in sequence. The function coercion $\cnffunc{c}{d}$
applies $c$ to the argument of a function and $d$ to the return
value.

Coercions come with a reduction relation so it is natural to ask about
their normal forms. \figref{fig:normal-coercion} defines the syntax of
coercion in normal form, written $\hat{c}$. Conceptually, a coercion
in normal form has three parts, all of which are optional.  It may
start with a projection $I?^l$, followed by a function coercion, and
then concluded with an injection $I!$ or a failure $\bot^l$. While
this is a simple idea, there are 10 clauses in the definition for
$\hat{c}$!

\section{A Framework for Proving Correctness of Cast ADTs}
\label{sec:framework}

% All cast representation shall be proved correct, that is, besides reducing 
% space consumption, using a cast representation should not affect the results of 
% program execution.
In this section we present a framework for proving the 
correctness of cast representations. The framework includes an abstract data 
type named Cast ADT, which specifies the operators of cast representations, 
a CEK machine $\SMachine{C}$, which uses the cast operators implemented by 
$C$ to work with casts, and theorems relating $\SMachine{C}$ to $\RMachine{B}$.
%
Subsection~\ref{sec:cast-adt-and-space-efficient-cek} describes Cast ADT and the 
machine \SMachine{C}. We conjecture that the machine \SMachine{C} is 
space-efficient provided that the cast representation $C$ performs compression.
%
Subsection~\ref{sec:good-cast-adt} defines two subset of Cast ADT, namely Lazy D 
Cast ADT and Lazy UD Cast ADT. As their names suggest, if $C$ is a Lazy D Cast 
ADT, executing a program in $\SMachine{C}$ would give the same result as in 
$\LDMachine$. Similarly, if $C$ is a Lazy UD Cast ADT, executing a program in 
$\SMachine{C}$ would give the same result as in $\LUDMachine$.
%
We prove the theorems relating \SMachine{C} and \RMachine{B} in
Subsection~\ref{sec:S-correct}.
The framework is written in the Agda language~\citep{norell2008dependently}.

%In Section~\ref{chap:background} we have seen that the standard dynamic 
%semantic of the cast calculi has a space-efficiency problem 
%(Section~\ref{sec:space-efficiency-problem}): casts might accumulate in 
%higher-order values and continuations and consume unbounded space. The 
%solution 
%is space-efficient cast representations, such as coercions in normal form and 
%threesomes.
%%
%Every such cast representations shall be proved correct. Actually, 
%people have done such proofs several times, such as the \emph{Correctness of 
%the threesome calculus} in \citep{Siek:2010:TWB:1706299.1706342} and the 
%bisimulation theorems in \citep{Siek:2015:BCT:2737924.2737968}.
%%
%These theorems are similar -- conceptually the only difference is the cast 
%representation.
%%
%But having two proofs blurs the similarity between the proofs and the key 
%differences between cast representations. 
%%
%Furthermore, there might be more cast representations in the future.
%Actually we introduce a new one in Section~\ref{chap:hypercoercions}.
%%
%Consequently, it is desirable to have generic proofs that handles many cast 
%representations that implement a common interface.
%
%Our solution is an abstract data type (Cast ADT), which specifies the 
%operators 
%of cast representations and a CEK machine ($\SMachine{C}$) parameterized over 
%Cast ADT (Section~\ref{sec:cast-adt-and-space-efficient-cek}). Then we 
%prove that if $C$ is a Lazy D Cast ADT then \SMachine{C} is equivalent to 
%\LDMachine, and that if $C$ is a Lazy UD Cast ADT then \SMachine{C} is 
%equivalent to \LUDMachine.
%%
%We conjecture that the machine \SMachine{C} is space-efficient provided that 
%the cast representation $C$ performs compression. We will prove or 
%conjecture that all cast representations defined 
%in the literature are instances of this ADT. 
%In \secref{sec:LazyD-correct} we prove that, for any
%instance $C$ of the Cast ADT, if $C$ satisfies a more refined abstract
%data type for Lazy D casts, then \SMachine{C} is equivalent to
%\LDMachine, that is, 
%\[
%\machineequiv{\LDMachine}{\SMachine{C}}
%\]
%In Section~\ref{sec:LazyUD-correct} we prove a corresponding \LUD\ theorem, 
%that is, for any instance $C$ of the Cast ADT, if $C$ satisfies a more refined 
%abstract data type for \LUD\ casts
%\[
%\machineequiv{\LUDMachine}{\SMachine{C}}
%\]
%
%Finally, we prove that some cast representations in 
%the literature (\LD\ and \LUD\ coercions in normal form, and \LUD\ threesomes) 
%are either instances of the Lazy D Cast ADT or the Lazy UD Cast ADT 
%(\secref{sec:coercions-correct} and \secref{sec:threesomes-correct}).
%We conjecture that supercoercions also fit into the two category.

\subsection{Cast Abstract Data Type and a Space-efficient CEK Machine}
\label{sec:cast-adt-and-space-efficient-cek}

\begin{figure}
	\[
	\begin{array}{lrcl}
	\stxrule{c,d}{Runtime casts}{\text{a data type defined by the 
	instance of Cast ADT}}
	\stxrule{v,u}{Values}{
		\vdyn{I}{v}  \mid
		\vtrue       \mid
		\vfalse                     
	}
	\stxrulecont{
		\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{c}{c}} \mid
		\ecast{\vcons{v}{v}}{\Pprod{c}{c}}
	}
	\stxrule{I,J}{Injectable types (\LD)}{P}
	\stxrule{I,J}{Injectable types (\LUD)}{G}
	\stxrule{E}{Environments}{
		\text{a partial function}\;\{\langle x , v \rangle , \dots \}
	}
	\stxrule{F}{Frames}{
		\FappXO{e}{E}  \mid
		\FappOX{v}     \mid
		\Fcar          \mid
		\Fcdr
	}
	\stxrulecont{
		\FconsXO{e}{E}  \mid
		\FconsOX{v}
	}
	\stxrule{K}{Continuations}{
		\Kcons{\Fcast{c}}{k}
	}
	\stxrule{k}{Pre-continuations}{
		\Knil         \mid
		\Kcons{F}{K}
	}
	\stxrule{s}{States}{
		\sexpr{e}{E}{K}  \mid
		\scont{v}{K}     \mid
		\shalt{v}        \mid
		\error{l}
	}
	\stxrule{r}{Cast results}{
		v \mid
		\error{l}
	}
	\end{array}
	\]

	\caption{Syntax of the space-efficient machine \SMachine{C} and auxiliary 
	data structures.}
	\label{fig:SEMachine-state}
\end{figure}

\figref{fig:SEMachine-state} defines the syntax of the space-efficient CEK 
machine (\SMachine{C}). This machine is parameterized over Cast ADT $C$.
%
The key differences between \SMachine{C} and the \RMachine{B} machine are the 
places where casts can accumulate~(\secref{sec:space-efficiency-problem}), that 
is, values and continuations.

Let $u,v$ range over values.
%
Recall that in \RMachine{B} we have two value constructors for functions and 
another two for pairs. That setup allows zero-or-more layers of casts at the 
top of higher-order values, which might lead to space inefficiency. 
%
In \SMachine{C}, we instead have exactly one layer of cast. 
Thus non-casted values in \RMachine{B} correspond to values 
in \SMachine{C} where casts are identities ($\mathit{id}(T)$). The 
$\mathit{id}$ functions is provided by the Cast ADT.

Let $K$ ranges over continuations and $k$ ranges over pre-continuations.
%
Pre-continuations are like the continuations in \RMachine{B}, but their 
top-most frame must not be a cast frames.
%
A continuation is now a pre-continuation prefixed with a cast frame.
%
Recall that continuations in \RMachine{B} have zero or more casts at the top.
In \SMachine{C}, however, every continuation has exactly one cast at the 
top.
%
Continuations in \RMachine{B} that have no casts at the top correspond
to continuations in \SMachine{C}\ whose casts are identities constructed by 
$\id{T}$.
Continuations in \RMachine{B} that have many casts at the top 
correspond
to continuations in \SMachine{C}\ where those casts are composed by
$\compose{c}{c}$, another functions provided by the Cast ADT.
%
Injectable types ($I,J$), environments ($\mathcal{E}$), states ($s$), and cast 
results ($r$) are the same as in \RBMachine.

Value typing (\valuetyping{v}{T}), continuation typing 
(\casttyping{k}{S}{T} and \casttyping{K}{S}{T}), and environment typing 
(\envtyping{\Gamma}{E}) are defined in the obvious way. 
They rely on cast typing \casttyping{c}{S}{T}, which shall be provided by the 
Cast~ADT.

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
	\label{def:cast-rep}
	A cast abstract data type is a set $Cast$, which is indexed by two types, 
	with four operators and two properties of operators:
\begin{itemize}
	\item $id(T)=c$: create an identity cast
	\item $\compose{c}{c}=c$: compose casts
	\item $\translate{\ccast{S}{l}{T}}=c$: translate cast in source 
		program to runtime casts
	\item $\denote{c}(v)=r$: apply a cast to a value
	\item If $ \valuetyping{v}{T} $ then $\denote{\id{T}}(v)=v$
	\item If $ \valuetyping{v}{T_1} $ and
		$ \casttyping{c}{T_1}{T_2} $ and
		$ \casttyping{d}{T_2}{T_3} $ 
		then 
		$ \denote{\compose{c}{d}}(v) = (\denote{c}(v)\mbind\denote{d})$
\end{itemize}
We use the syntax \casttyping{c}{S}{T} to mean $c$ is in the set 
$\textit{Cast}_{S,T}$.
\end{definition}

Before moving to the transition relation of \SMachine{C}, let's define two 
helper functions that manipulate continuations.
%
The function $\mathit{mkCont}$ construct a continuation from a pre-continuation 
by topping the pre-continuation with an identity cast. 
%
The function $\mathit{extCont}$ composes $c$ with the cast at the top of $K$ by 
using the composition operator 
$\fatsemi$.
\[
\begin{array}{rclr}
\figequalto{\mkCont{k}}{\Kcons{\Fcast{\id{S}}}{k}}{\sidecond{\casttyping{k}{S}{T}}}
\figequalto{\extCont{c}{\Kcons{\Fcast{d}}{k}}}{
	\Kcons{\Fcast{(\compose{c}{d})}}{k}
}{}
\end{array}
\]

\begin{figure}
	\fbox{\reducetoS{C}{s}{s}} Transition
	\[
	\begin{array}{rclr}
	\figreduceto{\sexpr{\evar{x}}{E}{K}}{\scont{E(x)}{K}}{}
	\figreducetoH{
		\sexpr{\elam{x}{S}{e}}{E}{K}
	}{
		\scont{\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{id(S)}{id(T)}}}{K}
	}{}
	& &
	\highlight{
	\sidecond{
		\expressiontyping{\Gamma,S}{e}{T}
		\text{ and }
		\Gamma \vdash E}
	}
	\\
	\figreduceto{\sexpr{\eapp{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\FappXO{e_2}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\etrue}{E}{K}}{\scont{\vtrue}{K}}{}
	\figreduceto{\sexpr{\efalse}{E}{K}}{\scont{\vfalse}{K}}{}
	\figreduceto{\sexpr{\eif{e_1}{e_2}{e_3}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\econs{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\FconsXO{e_2}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\ecar{e}}{E}{K}}{
		\sexpr{e}{E}{\mkCont{\Kcons{\Fcar}{K}}}
	}{}
	\figreduceto{\sexpr{\ecdr{e}}{E}{K}}{
		\sexpr{e}{E}{\mkCont{\Kcons{\Fcdr}{K}}}
	}{}
	\figreducetoH{
		\sexpr{\ecast{e}{\ccast{S}{l}{T}}}{E}{K}
	}{
		\sexpr{e}{E}{\extCont{\translate{\ccast{S}{l}{T}}}{K}}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\Fcast{c}}{k}}}{
		\denote{c}(v) \; \mbind \; \lambda u. \applyCont{u}{K}
	}{}
	\end{array}
	\]

	\fbox{$\applyCont{v}{k}=s$}
	\[
	\begin{array}{rclr}
	\figequalto{\applyCont{v_1}{\Kcons{\FappXO{e_2}{E}}{K}}}{\sexpr{e_2}{E}{
			\mkCont{\Kcons{\FappOX{v_1}}{K}}}
	}{}
	\figequaltoH{
		\applyCont{v}{\Kcons{\FappOX{(\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{c}{d}})}}{K}}
	}{
		\denote{c}(v) \mbind \lambda u. \
		\sexpr{e}{E[x:=u]}{\extCont{d}{K}}
	}{}
	\figequalto{\applyCont{\vtrue}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_2}{E}{K}}{}
	\figequalto{\applyCont{\vfalse}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_3}{E}{K}}{}
	\figequalto{\applyCont{v_1}{\Kcons{\FconsXO{e_2}{E}}{K}}}{
		\sexpr{e_2}{E}{\mkCont{\Kcons{\FconsOX{v_1}}{K}}}
	}{}
	\figequaltoH{\applyCont{v_2}{\Kcons{\FconsOX{v_1}}{K}}}{
		\scont{\ecast{\vcons{v_1}{v_2}}{\Pprod{\id{T_1}}{\id{T_2}}}}{K}
	}{}
	& & 
	\highlight{\sidecond{
%		\casttyping{K}{\Pprod{T_1}{T_2}}{T'}
		\valuetyping{v_1}{T_1}
		\text{ and }
		\valuetyping{v_2}{T_2}
	}}
	\\
	\figequaltoH{
		\applyCont{\ecast{\vcons{v_1}{v_2}}{\Pprod{c_1}{c_2}}}{\Kcons{\Fcar}{K}}
	}{
		\denote{c_1}(v_1) \mbind \lambda u.
		\scont{u}{K}
	}{}
	\figequaltoH{
		\applyCont{\ecast{\vcons{v_1}{v_2}}{\Pprod{c_1}{c_2}}}{\Kcons{\Fcdr}{K}}
	}{
		\denote{c_2}(v_2) \mbind \lambda u.
		\scont{u}{K}
	}{}
	\figequalto{\applyCont{v}{\Knil}}{\shalt{v}}{}
	\end{array}
	\]

	\fbox{\evalto{e}{o}} Evaluation
	\[
	\inference{
		\transitivelyreduceto{\sexpr{e}{\emptyset}{\mkCont{\Knil}}}{\shalt{v}}
	}{
		\evalto{e}{observe(v)}
	} \quad
	\infr{
		\transitivelyreduceto{\sexpr{e}{\emptyset}{\Knil}}{\error{l}}
	}{
		\evalto{e}{\error{l}}}
	\]
	\caption{Dynamic semantics of the cast calculi as a 
	space-efficient CEK	machine, \SMachine{C}. The transitions that involve 
	casts are highlighted
		in red.}
	\label{fig:semachine-dynamics}
\end{figure}

\figref{fig:semachine-dynamics} defines the transition relation 
$\reducetoS{C}{s}{s}$. They rely on functions provided 
by $C$ to work with casts.
%
To construct a function, the machine fills the casts with identities.
%
When evaluating a cast expression, the machine extends the continuation with 
cast \translate{\ccast{S}{l}{T}}.
%
To return a value to a continuation, the machine first applies the top cast to 
the value. If the cast application fails, the machine halts with the blame 
label from the failure. Otherwise, the machine handles the pre-continuation 
with the function $\applyCont{v}{k}$.
%
To apply a function, the machine applies the domain cast $c$ to the 
argument $v$ and evaluates the function body. The codomain cast $d$ is merged 
into the continuation with $\textit{extCont}$.
%
When the machine constructs a pair, it fills the casts with $id$s, just 
like how it did for functions. 
%
To take out the first part of a pair, the machine applies the first pending 
cast on the first value and returns the result to the continuation.
%
Taking out the second part of a pair is similar.

Reflexive transitive closure of reduction ($\reducetoS{C}{s}{s}$) and
evaluation ($\evaltoS{C}{e}{o}$) are the same as in \RMachine{B}.

\begin{proposition}[\SMachine{C} is deterministic]
	\label{prop:SEMachine-deterministic}
	If $\reducetoS{C}{s}{s_1}$ and $\reducetoS{C}{s}{s_2}$ 
	then $s_1 = s_2$.
\end{proposition}

\subsection{Specifying good instances of Cast ADT}
\label{sec:good-cast-adt}

Ultimately, we want to claim that if a $C$ satisfies a collection of 
properties, \SMachine{C} would be equivalent to \LDMachine\ or \LUDMachine. 
Cast ADT has specified the behavior of identities and composition, so the only 
remaining operator is translation.
%
Lazy D Cast ADT (Definition~\ref{def:LazyD-CastADT}) and Lazy UD Cast ADT 
(Definition~\ref{def:LazyUD-CastADT}) specify the behavior 
of \translate{\ccast{S}{l}{T}}. Readers are encouraged to compare them with the 
corresponding $\mathit{applyCast}$ functions 
(Fig.~\ref{fig:blame-strategy-LazyD} and 
Fig.~\ref{fig:blame-strategy-LazyUD}).

\begin{definition}[Lazy D Cast ADT]
	\label{def:LazyD-CastADT}
	A Cast is a Lazy D Cast if:
	\begin{enumerate}
		\item $\denote{\translate{\ccast{\Tdyn}{l}{\Tdyn}}}(v) = v$
		\item $\denote{\translate{\ccast{P}{l}{\Tdyn}}}(v) = \vdyn{P}{v}$
		\item $\denote{\translate{\ccast{\Tdyn}{l}{Q}}}(\vdyn{P}{u}) = 
		\denote{\translate{\ccast{P}{l}{Q}}}(u)$
		\item $\denote{\translate{\ccast{\Pbool}{l}{\Pbool}}}(v) = v$
		\item 
		$\denote{\translate{\ccast{\Pfunc{S_1}{T_1}}{l}{\Pfunc{S_2}{T_2}}}}(
			\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{c}{d}})$\\
		$=
		\ecast{\vfunc{x}{S}{e}{E}}{
			\Pfunc{
				(\compose{\translate{\ccast{S_2}{l}{S_1}}}{c})}{
				(\compose{d}{\translate{\ccast{T_1}{l}{T_2}}})}}$
		\item 
		$\denote{\translate{\ccast{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}(
		\ecast{\vcons{v_1}{v_2}}{\Pprod{c}{d}})$\\
		$=
		\ecast{\vcons{v_1}{v_2}}{
			\Pprod{
				(\compose{c}{\translate{\ccast{S_1}{l}{S_2}}})}{
				(\compose{d}{\translate{\ccast{T_1}{l}{T_2}}})}}$
		\item $\denote{\translate{\ccast{P}{l}{Q}}}(v) = \error{l}$\\
			if $\notshallowlyconsistent{P}{Q}$
	\end{enumerate}
\end{definition}

\begin{definition}[Lazy UD Cast ADT]
	\label{def:LazyUD-CastADT}
	A Cast is a Lazy UD Cast if:
	\begin{enumerate}
	\item $\denote{\translate{\ccast{\Tdyn}{l}{\Tdyn}}}(v) = v$
	\item $\denote{\translate{\ccast{G}{l}{\Tdyn}}}(v) = \vdyn{G}{v}$
	\item $\denote{\translate{\ccast{P}{l}{\Tdyn}}}(v) = 
		\denote{\translate{\ccast{P}{l}{G}}}(v) \mbind 
		\denote{\ccast{G}{l}{\Tdyn}}$\\
		if $G \sim P$ and $G \neq P$	
	\item $\denote{\translate{\ccast{\Tdyn}{l}{G}}}(\vdyn{G}{u}) = u$
	\item $\denote{\translate{\ccast{\Tdyn}{l}{H}}}(\vdyn{G}{u}) = \error{l}$\\
		if $G \neq H$
	\item $\denote{\translate{\ccast{\Tdyn}{l}{Q}}}(v) = 
		\denote{\translate{\ccast{\Tdyn}{l}{H}}}(v) \mbind
		\denote{\translate{\ccast{H}{l}{Q}}}$\\
		if $H \sim Q$ and $H \neq Q$
	\item $\denote{\translate{\ccast{\Pbool}{l}{\Pbool}}}(v) = v$
		\item 
		$\denote{\translate{\ccast{\Pfunc{S_1}{T_1}}{l}{\Pfunc{S_2}{T_2}}}}(
		\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{c}{d}})$\\
		$=
		\ecast{\vfunc{x}{S}{e}{E}}{
			\Pfunc{
				(\compose{\translate{\ccast{S_2}{l}{S_1}}}{c})}{
				(\compose{d}{\translate{\ccast{T_1}{l}{T_2}}})}}$
		\item 
		$\denote{\translate{\ccast{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}(
		\ecast{\vcons{v_1}{v_2}}{\Pprod{c}{d}})$\\
		$=
		\ecast{\vcons{v_1}{v_2}}{
			\Pprod{
				(\compose{c}{\translate{\ccast{S_1}{l}{S_2}}})}{
				(\compose{d}{\translate{\ccast{T_1}{l}{T_2}}})}}$
		\item $\denote{\translate{\ccast{P}{l}{Q}}}(v) = \error{l}$\\
			if $\notshallowlyconsistent{P}{Q}$
	\end{enumerate}
\end{definition}

\subsection{The Space-efficient CEK Machine is Correct}
\label{sec:S-correct}

In this section, we prove that if $C$ is a Lazy D Cast ADT then
\[
\machineequiv{\LDMachine}{\SMachine{C}}
\]
and if $C$ is a Lazy UD Cast ADT then
\[
\machineequiv{\LUDMachine}{\SMachine{C}}
\]
The proof is done via a weak bisimulation.

\begin{figure}

\fbox{\bisimc{\overline{\ccast{S}{l}{T}}}{c}{S}{T}}
\begin{gather*}
\infr{}{\bisimc{[]}{id(T)}{T}{T}} \quad
\infr{}{\bisimc{[\ccast{S}{l}{T}]}{\translate{\ccast{S}{l}{T}}}{S}{T}}
\\[1ex]
%\infr{
%	\bisimc{z_1}{c_1}{T_1}{T_2}\\
%	\bisimc{z_2}{c_2}{T_2}{T_3}
%}{
%	\bisimc{(z_1 + z_2)}{(\compose{c_1}{c_2})}{T_1}{T_n}
%}
\infr{
	\bisimc{
		(\ccast{T_1}{l_1}{T_2}\ldots\ccast{T_{i-1}}{l_{i-1}}{T_i})
	}{c_1}{T_1}{T_i}\\
	\bisimc{
		(\ccast{T_i}{l_i}{T_{i+1}}\ldots\ccast{T_{n-1}}{l_{n-1}}{T_n})
	}{c_2}{T_i}{T_n}
}{
	\bisimc{
		(\ccast{T_1}{l_1}{T_{2}}\ldots\ccast{T_{n-1}}{l_{n-1}}{T_n})
	}{
		(\compose{c_1}{c_2})
	}{T_1}{T_n}
}
\end{gather*}

\fbox{$\bisimK{K}{K}{S}{T}$}
\begin{gather*}
\infr{
	\bisimK{K}{k'}{S_n}{T} \qquad
	\bisimc{
		(\ccast{S_1}{l_1}{S_2} \ldots \ccast{S_{n-1}}{l_{n-1}}{S_n})
	}{c}{S_1}{S_n}
}{
	\bisimK{
		\Kcons{
			\Fcast{\ccast{S_1}{l_1}{S_2}}
		}{\ldots\Kcons{\Fcast{\ccast{S_{n-1}}{l_{n-1}}{S_n}}}{K}}
	}{
		\Kcons{\Fcast{c'}}{k'}
	}{S_1}{T}
}
\end{gather*}

\fbox{$\bisimv{v}{v}{T}$}
\begin{gather*}
\infr{
	\expressiontyping{\Gamma,x : S_1}{e}{T_1}
	\qquad
	\bisimE{\Gamma}{E}{E'}
	\\
	\bisimc{
		(\ccast{S_n}{l_{n-1}}{S_{n-1}} \ldots
		\ccast{S_2}{l_1}{S_1})
	}{c}{S_n}{S_1}
	\\
	\bisimc{
		(		\ccast{T_1}{l_{1}}{T_2} \ldots
		\ccast{T_{n-1}}{l_{n-1}}{T_n}
		)	}{d}{T_1}{T_n}
	\\
	v = \vfuncc{
		\vfuncc{
			\vfunc{x}{S_1}{e}{E}
		}{S_1}{T_1}{l_1}{S_2}{T_2}
		\ldots
	}{S_{n-1}}{T_{n-1}}{l_{n-1}}{S_n}{T_n}
}{
	\bisimv{v}{
		(\ecast{\vfunc{x}{S_1}{e}{E'}}{\Pfunc{c}{d}})
	}{
		\Pfunc{S_n}{T_n}}
}
\\[1ex]
\infr{
	\bisimv{v_1}{v_1'}{S_1} \qquad
	\bisimv{v_2}{v_2'}{T_1} \\
	\bisimc{
		(\ccast{S_1}{l_{1}}{S_2} \ldots
		\ccast{S_{n-1}}{l_{n-1}}{S_n})
	}{c}{S_1}{S_n} \\
	\bisimc{
		(\ccast{T_1}{l_{1}}{T_2} \ldots
		\ccast{T_{n-1}}{l_{n-1}}{T_n})
	}{d}{T_1}{T_n}
	\\ v = \vconsc{
		\vconsc{
			\vcons{v_1}{u_1}
		}{S_1}{T_1}{l_1}{S_2}{T_2}
		\ldots
	}{S_{n-1}}{T_{n-1}}{l_{n-1}}{S_n}{T_n}
}{
	\bisimv{
		v
	}{
		(\ecast{\vcons{v_2}{u_2}}{\Pprod{c}{d}})
	}{
		\Pprod{S_n}{T_n}}
}
\end{gather*}

\fbox{\bisimE{\Gamma}{E}{E}}
\begin{gather*}
\inference{
}{
	\bisimE{\emptyset}{\emptyset}{\emptyset}
}
\quad
\inference{
	\Gamma \vdash E \approx E' \qquad
	v \approx v' : T
}{
	\Gamma , x : T \vdash 
	E [ x := v' ] \approx E' [ x := v' ]
}
\end{gather*}

\fbox{\bisims{s}{s}{T}}
\begin{gather*}
\infr{}{\bisims{\error{l}}{\error{l}}{T}} \quad
\infr{
	\bisimv{v}{v'}{T}
}{
	\bisims{\shalt{v}}{\shalt{v'}}{T}
}
\\
\infr{
	\bisimv{v}{v'}{S} \\
	\bisimK{K}{K'}{S}{T}
}{
	\bisims{\scont{v}{K}}{\scont{v}{K'}}{T}
} \quad
\infr{
	\expressiontyping{\Gamma}{e}{S} \\
	\bisimE{\Gamma}{E}{E'}          \\
	\bisimK{K}{K'}{S}{T}
}{
	\bisims{\sexpr{e}{E}{K}}{\sexpr{e}{E'}{K'}}{T}
}
\end{gather*}

\caption{Bisimulation between \RBMachine\ and \SMachine{C}. It is the smallest 
congruence relation containing the shown rules. Everything to the left (resp. 
right) of the 
symbol $\approx$ is in \RBMachine\ (resp. \SMachine{C}).}
\label{fig:bisim-SC-D}
\end{figure}

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation between \RBMachine\ 
and \SMachine{C}. It is the smallest congruence relation containing the rules 
in Fig.~\ref{fig:bisim-SC-D}.
%
The relation $\bisimc{\overline{\ccast{S}{l}{T}}}{c}{S}{T}$ relates a sequence 
of cast in \RBMachine\ with one cast in \SMachine{C}. This relation is designed 
to cover all possibly ways to construct a cast representation. The empty 
sequence is related to the identity cast. A \RBMachine\ cast is related 
to its translation. Finally, cast sequencing preserves the bisimulation 
relation.
%
We remark that this bisimulation relation is unusual in that it uses
some ADT operations in its definition, such as $\id{T}$, 
$\translate{\ccast{S}{l}{T}}$, and $\compose{c}{c}$, which is necessary because 
\SMachine{C} and the bisimulation is parameterized over the Cast ADT.

The bisimulation relation is mutually defined on continuations and pre-continuations,
written 
$\bisimK{K}{K}{S}{T}$ and $\bisimK{K}{k}{S}{T}$, where $S$ and $T$ are their 
source and target types respectively, values, written $\bisimv{v}{v}{T}$, 
where $T$ is their type, and environments, written $\bisimE{\Gamma}{E}{E}$, 
where $\Gamma$ is their type environment. The bisimulation relation for values 
is designed to make Lemma~\ref{lem:LazyD-apply-cast-preserve-bisim} and 
Lemma~\ref{lem:LazyUD-apply-cast-preserve-bisim} (described below) true, that 
is, if $\bisimv{v}{v'}{S}$
then
\[
\bisimr{
	\applyCast{v}{\ccast{S}{l}{T}}
}{
	\denote{\translate{\ccast{S}{l}{T}}}(v')
}{T}
\]
So if two values of type $I$ are related, injecting them to $\Tdyn$
gives related values.
%
To relate function closures, the domain and codomain casts should be related. 
Closures in \RMachine{B} are proxied with a sequence of function casts, which 
can be viewed as two cast sequences in the obvious way.
%
The rule for products is similar.

One nice aspect of using abstract machines for this proof is that
the expressions in the machine are simply related by syntactic equality.

\begin{lemma}[Lazy D $\mathit{applyCast}$ Preserves Bisimulation]
	\label{lem:LazyD-apply-cast-preserve-bisim}
	Assume $C$ is a Lazy D Cast ADT,
	$v$ is a value in \LDMachine,
	$v'$ is a value in \SMachine{C},
	and $\bisimv{v}{v'}{S}$
	\[
	\bisimr{
		\applyCast{v}{\ccast{S}{l}{T}}
	}{
		\denote{\translate{\ccast{S}{l}{T}}}(v')
	}{T}
	\]
\end{lemma}
\begin{proof}
	Case splitting whether $S$ is shallowly consistent with $T$.
	If yes, case splitting how they are shallowly consistent.
	Then apply properties of Lazy D Cast when applicable.
\end{proof}

\begin{lemma}[Lazy UD $\mathit{applyCast}$ Preserves Bisimulation]
	\label{lem:LazyUD-apply-cast-preserve-bisim}
	Assume $C$ implements Lazy UD Cast ADT,
	$v$ is a value in \LUDMachine,
	$v'$ is a value in \SMachine{C},
	and $\bisimv{v}{v'}{S}$
	\[
	\bisimr{
		\applyCast{v}{\ccast{S}{l}{T}}
	}{
		\denote{\translate{\ccast{S}{l}{T}}}(v')
	}{T}
	\]
\end{lemma}
\begin{proof}
	Case splitting whether $S$ is shallowly consistent with $T$.
	If yes, case splitting how they are shallowly consistent.
	Then apply properties of Lazy UD Cast when applicable.
\end{proof}
%
%
%Before moving to the bisimulation proof, let's introduce two lemmas that 
%manipulate continuations.
%\begin{lemma}[$\mathit{mkCont}$ Preserves Bisimulation]
%	If \bisimK{K}{k}{S}{T} then
%	\[
%	\bisimK{K}{\mkCont{k}}{S}{T}.
%	\]
%\end{lemma}
%\begin{lemma}[$\mathit{extCont}$ Preserves Bisimulation]
%	If $\bisimc{
%		(\ccast{S_1}{l_1}{S_2}\ldots\ccast{S_{n-1}}{l_{n-1}}{S_n})
%	}{c}{S_1}{S_n}$
%	and $\bisimK{K}{K'}{S_n}{T}$ then
%	\[
%	\bisimK{
%		\Kcons{
%			\Fcast{\ccast{S_1}{l_1}{S_2}}
%		}{\ldots\Kcons{\Fcast{\ccast{S_{n-1}}{l_{n-1}}{S_n}}}{K}}
%	}{
%		\extCont{c}{K'}
%	}{S_1}{T}
%	\]
%\end{lemma}

We now come to the main lemmas, the weak bisimulation between
\RBMachine\ and \SMachine{C}.

\begin{lemma}[Lazy D Weak Bisimulation between \LDMachine\ and \SMachine{C}]
	\label{lem:weak-bisim-SC-D}
	Assume $C$ implements Lazy D Cast ADT
	and $s_1$ is a state in \LDMachine\
	and $s_1'$ is a state in \SMachine{C}
	and $\bisims{s_1}{s_1'}{T}$
	then either
\begin{enumerate}
\item $s_1  = \error{l}$
  and $s_1' = \error{l}$
  for some $l$, or
\item $s_1 = \shalt{v}$
  and $s_1' = \shalt{v'}$
  and $\bisimv{v}{v'}{T}$
  for some $v$ and $v'$, or
\item $\reduceto{s_1}{^{+}s_2}$
  and $\reduceto{s_1'}{^{+}s_2'}$
  and $\bisims{s_2}{s_2'}{T}$
\end{enumerate}
\end{lemma}
%\begin{proof}
%We proceed by case analysis on $\bisims{s_1}{s_1'}{T}$. 
%\begin{description}
%\item[Case $\inference{
%}{
%	\bisims{\error{l}}{\error{l}}{T}
%}$]
%This is case (1).
%\item[Case $\inference{
%	\bisimv{v}{v'}{T}
%}{
%	\bisims{\shalt{v}}{\shalt{v'}}{T}
%}$]
%This is case (2).
%\item[Case $\inference{
%	\expressiontyping{\Gamma}{e}{T} \\
%	\bisimE{\Gamma}{E}{E'} \\
%	K \approx K' : T
%}{
%	\bisims{\sexpr{e}{E}{K}}{\sexpr{e}{E'}{K'}}{T}
%}$]
%This is case (3). By case analysis on $e$.
%\item[Case $\inference{
%	\bisimv{v}{v'}{S} \\
%	\bisimK{K}{K'}{S}{T}
%}{
%	\bisims{\scont{v}{K}}{\scont{v'}{K'}}{T}
%}$]
%This is case (3). By inversion on $\bisimK{K}{K'}{S}{T}$, we have
%\[
%\begin{array}{rcl}
%K  & = &
%\Kcons{
%	\Fcast{\ccast{S_1}{l_1}{S_2}}
%}{\ldots\Kcons{\Fcast{\ccast{S_{n-1}}{l_{n-1}}{S}}}{K_1}} \\
%K' & = &
%\Kcons{\Fcast{c'}}{k_1'}
%\end{array}
%\]
%After zero or more transition steps, $\scont{v}{K}$ will either transition to 
%$\scont{u}{K_1}$ for some $u$ or transition to $\error{l}$ for some $l$ if one 
%of the cast fail. On the \SMachine{C} side, $\scont{v'}{K'}$ will either 
%transition to $\applyCont{u'}{k_1'}$ for some $u'$ or transition to 
%$\error{l'}$ for some $l'$ if the cast fail. Because the sequence of casts in 
%$K$ bisimulates the cast in $K'$, we can show with  
%Lemma~\ref{lem:LazyD-apply-cast-preserve-bisim} that either
%\begin{itemize}
%	\item all casts succeed and	$\bisimv{u}{u'}{S}$, or
%	\item both machines transition to error state and $l = l'$
%\end{itemize}
%When all casts succeeded, we also have $\bisimK{K_1}{k_1'}{S}{T}$ and need to 
%show that $\reduceto{\scont{u}{K_1}}{s_2}$ and $\applyCont{u'}{k_1'} = s_2'$ 
%and $\bisims{s_2}{s_2'}{T}$ for some $s_2$ and $s_2'$.
%%
%We proceed by case analysis on $K$. The most interesting cases 
%are function application and pair projection. Due to space limitation, 
%we only illustrate the general idea with function applications
%
%\begin{description}
%\item[Subcase $\infr{
%	\bisimv{v_1}{v_1'}{\Pfunc{S}{T_1}} \qquad
%	\bisimK{K_2}{K_2'}{T_1}{T}
%}{
%	\bisimK{
%		\Kcons{\FappOX{v_1}}{K_2}
%	}{
%		\Kcons{\FappOX{v_1'}}{K_2'}
%	}{S}{T}
%}$] By inversion on $\bisimv{v_1}{v_1'}{\Pfunc{S}{T_1}}$ we know the proxies 
%in $v_1$ bisimulate the proxies in $v_1'$. After a reasoning similar to the 
%above one that we used to reason about casts in continuations, we know that 
%either both machines will transition to error states if the input proxies fail 
%in casting the argument, or both machines will transition to a state 
%evaluation 
%the function body.
%\todo{need to elaborate this part, or just delete it.}
%\end{description}
%\end{description}
%\end{proof}

\begin{lemma}[Lazy UD Weak Bisimulation between \LUDMachine\ and \SMachine{C}]
	\label{lem:weak-bisim-SC-UD}
	Assume $C$ implements Lazy UD Cast ADT
	and $\bisims{s_1}{s_1'}{T}$
	then either
	\begin{enumerate}
		\item $s_1  = \error{l}$
		and $s_1' = \error{l}$
		for some $l$, or
		\item $s_1 = \shalt{v}$
		and $s_1' = \shalt{v'}$
		and $\bisimv{v}{v'}{T}$
		for some $v$ and $v'$, or
		\item $\reduceto{s_1}{^{+}s_2}$
		and $\reduceto{s_1'}{^{+}s_2'}$
		and $\bisims{s_2}{s_2'}{T}$
	\end{enumerate}
\end{lemma}
%\begin{proof}
%	Similar to Lemma~\ref{lem:weak-bisim-SC-D}.
%\end{proof}

\begin{theorem}[Correctness of Lazy D \SMachine{C}]
	\label{thm:SD-equiv}
	Suppose $C$ is a Lazy D Cast ADT
	\[
	\machineequiv{\RBMachine}{\SMachine{C}}
	\]
\end{theorem}
\begin{proof}
	By the definitions of $\textit{eval}$, we need to show 
	\[
	\transitivelyreducetoM{\RMachine{B}}{\sexpr{e}{\emptyset}{\Knil}}{
		s
	}
	\;\text{if and only if}\;
	\transitivelyreducetoM{\SMachine{C}}{\sexpr{e}{\emptyset}{\mkCont{\Knil}}}{
		s'
	}
	\]
	where $s$ and $s'$ are bisimuated final states, i.e. of the form 
	$\shalt{v}$ or $\error{l}$. The initial states are related by the 
	bisimulation relation, so we can prove this theorem with 
	Lemma~\ref{lem:weak-bisim-SC-D}.
\end{proof}

\begin{theorem}[Correctness of Lazy UD \SMachine{C}]
	\label{thm:SUD-equiv}
	Suppose $C$ is a Lazy UD Cast ADT
	\[
	\machineequiv{\RBMachine}{\SMachine{C}}
	\]
\end{theorem}
\begin{proof}
	Similar to Theorem~\ref{thm:SD-equiv}.
\end{proof}

\section{Hypercoercions, a New Cast Representation} \label{sec:hypercoercions}

\begin{figure}
	Syntax of Hypercoercions
	\[
	\begin{array}{lrcl}
	\stxrule{c}{Hypercoercions}{
		\hcci           \mid
		\hccc{h}{m}{t}
	}
	\stxrule{h}{Heads}{
		\hche     \mid
		\hchp{I}{l}
	}
	\stxrule{m}{Middles}{
		\hcmbool       \mid
		\hcmfunc{c}{c} \mid
		\hcmprod{c}{c} \mid
		\hcmfail{l}    
	}
	\stxrule{t}{Tails}{
		\hcte     \mid
		\hcti{I}
	}
	\end{array}
	\]

	\fbox{$\mathit{id}( T ) = c$}
	\[
	\begin{array}{rclr}
	\figequalto{id(\star)}{
		\hcci
	}{}
	\figequalto{id(P)}{
		\hccc{\hche}{id(P)}{\hcte}
	}{}
	\end{array}
	\]

	\fbox{$\mathit{id}_m( P ) = m$}
	\[
	\begin{array}{rclr}
	\figequalto{id(\Pbool)}{\hcmbool}{}
	\figequalto{id(\Pfunc{S}{T})}{
		\hcmfunc{id(S)}{id(T)}
	}{}
	\figequalto{id(\Pprod{S}{T})}{
		\hcmprod{id(S)}{id(T)}
	}{}
	\end{array}
	\]

	\fbox{\casttyping{c}{S}{T}} Hypercoercion typing 
	\begin{gather*}
	\inference{}{\casttyping{\hcci}{\Tdyn}{\Tdyn}}
	\quad
	\inference{
		\casttyping{h}{S}{P} &
		\casttyping{m}{P}{Q} &
		\casttyping{t}{Q}{T}
	}{
		\casttyping{\hccc{h}{m}{t}}{S}{T}
	}
	\end{gather*}

	\fbox{\casttyping{h}{T}{P}} Head typing
	\begin{gather*}
	\inference{}{\casttyping{\hche}{P}{P}}
	\quad
	\inference{}{\casttyping{\hchp{I}{l}}{\Tdyn}{I}}
	\end{gather*}

	\fbox{$\casttyping{m}{T}{T}$} Middle typing
	\begin{gather*}
	\infr{}{\casttyping{\hcmfail{l}}{P}{Q}} \quad
	\inference{}{\casttyping{\hcmbool}{\Pbool}{\Pbool}} \\
	\inference{
		\casttyping{c}{S_2}{S_1} &
		\casttyping{d}{T_1}{T_2}
	}{
		\casttyping{\hcmfunc{c}{d}}{\Pfunc{S_1}{T_1}}{\Pfunc{S_2}{T_2}}
	}
	\quad
	\inference{
		\casttyping{c}{S_1}{S_2} &
		\casttyping{d}{T_1}{T_2}
	}{
		\casttyping{\hcmprod{c}{d}}{\Pprod{S_1}{T_1}}{\Pprod{S_2}{T_2}}
	}
	\end{gather*}

	\fbox{\casttyping{t}{P}{T}} Tail typing
	\begin{gather*}
	\inference{}{\casttyping{\hcte}{P}{P}} \quad
	\inference{}{\casttyping{\hcti{I}}{I}{\Tdyn}} \quad
	\end{gather*}

	\caption{Definition of hypercoercions}
	\label{fig:hypercoercion}
\end{figure}

In this section we present a new cast representation called hypercoercions. It 
comes with two flavors: Lazy D hypercoercions and Lazy UD one.  The design of 
hypercoercions is motivated by the observation that a normal coercion has at most three
parts. Hypercoercions make this structure explicit: a hypercoercion $c$ is
either \hcci, the identity cast for \Tdyn, or it contains
three parts: a head, middle, and tail, as defined in
\figref{fig:hypercoercion}.
\begin{itemize}
	\item a head is either a projection or the no-op;
	\item a middle $m$ involves coercions that preserve a type
	constructor, i.e., coercions between Boolean, function, and pair, and 
	failure coercions;
	\item a tail is either an injection or the no-op.
\end{itemize}

Subsection~\ref{sec:LazyD-Hypercoercions} implements the \CastADT\ with the \LD\ 
hypercoercions and proves that \LD\ hypercoercions is a Lazy D Cast ADT. 
Subsection~\ref{sec:LazyUD-Hypercoercions} does the \LUD\ counterparts.
Finally, we argue in Subsection~\ref{sec:hypercoercions-compact} that 
hypercoercions would consume less space than coercions in 
normal form and hence are more suitable for implementation.

\subsection{\LD\ Hypercoercions}
\label{sec:LazyD-Hypercoercions}

\begin{figure} %% composition 1
\[
\begin{array}{lrcl}
\stxrule{g}{Gaps}{
	\gnone           \mid
	\gsome{S}{l}{T}
}
\end{array}
\]

\fbox{$\compose{c}{c} = c$} Composition of \LD\ hypercoercions
\[
\begin{array}{rclr}
\figequalto{\compose{c_1}{c_2}}{\composegap{c_1}{\gnone}{c_2}}{}
\end{array}
\]

\fbox{$\composegap{c}{g}{c} = c$}
\[
\begin{array}{rcll}
\figequalto{\composegap{\hcci}{g}{\hcci}}{\hcci}{}
\figequalto{\composegap{\hcci}{g}{\hccc{h}{m}{t}}}{
	\hccc{\mkhead{g}{h}}{m}{t}
}{}
\figequalto{\composegap{\hccc{h}{m}{t}}{g}{\hcci}}{
	\hccc{h}{m}{\mktail{t}{g}}
}{}
\figequalto{\composegap{\hccc{h_1}{m_1}{t_1}}{g}{\hccc{h_2}{m_2}{t_2}}}{
	\hccc{h_1}{(\composegap{m_1}{\mkgap{t_1}{g}{h_2}}{m_2})}{t_2}
}{}
\end{array}
\]

\fbox{\mkhead{g}{h}}
\[
\begin{array}{rclr}
\figequalto{\mkhead{g}{\hchp{P}{l}}}{\hchp{P}{l}}{}
\figequalto{\mkhead{\gsome{\Tdyn}{l}{P}}{\hche}}{\hchp{P}{l}}{}
%\mathit{mkProj(}
%& g
%& ,
%& \hchp{P}{l}
%& )=
%& \hchp{P}{l} \\
%\mathit{mkProj(}
%& \gsome{\Tdyn}{l}{P}
%& ,
%& \hche
%& )=
%& \hchp{P}{l}
\end{array}
\]
%\[
%\begin{array}{rcll}
%\figequalto{\mkhead{g}{\hchp{P}{l}}}{\hchp{P}{l}}{}
%\figequalto{\mkhead{\gsome{\Tdyn}{l}{P}}{\hche}}{\hchp{P}{l}}{}
%\end{array}
%\]

\fbox{\mktail{t}{g}}
\[
\begin{array}{rcll}
\figequalto{\mktail{\hcti{P}}{g}}{\hcti{P}}{}
\figequalto{\mktail{\hcte}{\gsome{P}{l}{\Tdyn}}}{\hcti{P}}{}
\end{array}
\]

\fbox{\mkgap{t}{g}{h}}
\[
\begin{array}{rcll}
\figequalto{\mkgap{\hcte}{g}{\hche}}{g}{}
\figequalto{\mkgap{\hcte}{\gsome{P}{l'}{\Tdyn}}{\hchp{Q}{l}}}{
	\gsome{P}{l}{Q}
}{}
\figequalto{\mkgap{\hcti{P}}{g}{\hchp{Q}{l}}}{
	\gsome{P}{l}{Q}
}{}
\figequalto{\mkgap{\hcti{P}}{\gsome{\Tdyn}{l}{Q}}{\hche}}{
	\gsome{P}{l}{Q}
}{}
\end{array}
\]

\fbox{$\composegap{m}{g}{m} = m$}
\[ 
\begin{array}{rclr}
\figequalto{\composegap{\hcmfail{l}}{g}{m_2}}{
	\hcmfail{l}
}{}
\figequalto{\composegap{m_1}{\gsome{P}{l}{Q}}{m_2}}{
	\hcmfail{l}
}{\sidecond{
		\notshallowlyconsistent{P}{Q}
		\text{ and } \forall l. m_1 \neq \hcmfail{l}
}}
\figequalto{\composegap{m_1}{g}{\hcmfail{l}}}{
	\hcmfail{l}
}{\sidecond{
		\goodgap{g}
		\text{ and } \forall l. m_1 \neq \hcmfail{l}}}
\figequalto{\composegap{\hcmbool}{g}{\hcmbool}}{
	\hcmbool
}{}
\figequalto{
	\composegap{
		(\hcmfunc{c_1}{d_1})}{
		g}{
		(\hcmfunc{c_2}{d_2})}
}{
	\hcmfunc{
		(\composegap{c_2}{\dom{g}}{c_1})}{
		(\composegap{d_1}{\cod{g}}{d_2})}
}{}
\figequalto{
	\composegap{
		(\hcmprod{c_1}{d_1})}{
		g}{
		(\hcmprod{c_2}{d_2})}
}{
	\hcmprod{
		(\composegap{c_1}{\car{g}}{c_2})}{
		(\composegap{d_1}{\cdr{g}}{d_2})}
}{}
\end{array}
\]

\fbox{$\goodgap{g}$}
\begin{gather*}
\infr{}{\goodgap{\gnone}} \quad
\infr{\shallowlyconsistent{P}{Q}}{\goodgap{\gsome{P}{l}{Q}}}
\end{gather*}

\fbox{$\dom{g}=g$}
\[
\begin{array}{rclr}
\figequalto{\dom{\gnone}}{\gnone}{}
\figequalto{\dom{\gsome{\Pfunc{S_1}{T_1}}{l}{\Pfunc{S_2}{T_2}}}}{
	\gsome{S_2}{l}{S_1}
}{}
\end{array}
\]

\fbox{$\cod{g}=g$}
\[
\begin{array}{rclr}
\figequalto{\cod{\gnone}}{\gnone}{}
\figequalto{\cod{\gsome{\Pfunc{S_1}{T_1}}{l}{\Pfunc{S_2}{T_2}}}}{
	\gsome{T_1}{l}{T_2}
}{}
\end{array}
\]

\fbox{$\car{g}=g$}
\[
\begin{array}{rclr}
\figequalto{\car{\gnone}}{\gnone}{}
\figequalto{\car{\gsome{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}{
	\gsome{S_1}{l}{S_2}
}{}
\end{array}
\]

\fbox{$\cdr{g}=g$}
\[
\begin{array}{rclr}
\figequalto{\cdr{\gnone}}{\gnone}{}
\figequalto{\cdr{\gsome{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}{
	\gsome{T_1}{l}{T_2}
}{}
\end{array}
\]

	\caption{Composition of \LD\ Hypercoercions}
	\label{fig:LazyD-Hypercoercions-composition}
\end{figure}

\figref{fig:LazyD-Hypercoercions-composition} defines the functions 
\compose{c}{c} that compose \LD\ hypercoercions. We use these functions to 
implement the \CastADT.

\todo[inline]{KC: Should we mention KC and Deyaa as citation?
(something like in private communication with KC and Deyaa)}

The \compose{c}{c} operator is the key to compressing coercions and maintaining 
space-efficiency. The composition operator for Coercions in Normal Form 
requires the target type of the first coercion to match the source type of the 
second. This requirement, however, imposes a considerable difficulty in 
termination proof of the composition of \LD\ Coercions in Normal Form, reported 
by Kuang-Chen Lu and Deyaaeldeen Almahallawi. So it is useful to relax this 
restriction for the \LD\ tracking strategy.
%
We shall need optional source cast, written $g$, that range over
$\gnone$ or $\gsome{S}{l}{T}$, to maintain the gap between the hypercoercions 
to be composed. The source cast is mandatory when the target
type of the first hypercoercion does not match the source type of the second. 
Then we write the composition of hypercoercions as \composegap{c_1}{g}{c_2}.
%  
When both $c_1$ and $c_2$ are \hcci, their composition is also \hcci. 
%
When the first hypercoercion is the \hcci, the output must start with a 
projection, in which case we invoke the helper function \textit{mkProj}. In 
\textit{mkProj}, if the head of $c_2$ is a projection, the result is $c_2$.  
Otherwise, we need a label to build the projection. Since the head of $c_2$ is 
the no-op, its source type must be a pretype. Thus we know $g$ must be a 
source cast and we use it to construct the projection.
%
Going back to $\composegap{c}{g}{c}$, when the second hypercoercion is the 
\hcci, the output must end with an injection, in which case we invoke the 
helper function \textit{mkInj}.
%
Finally, when both inputs are of the form \hccc{h}{m}{t}, we compose the middle 
with an updated gap.
%
The second and third cases of \mkgap{t}{g}{h} is interesting, where we choose 
the label in projection to be responsible for potential inconsistency.

Another interesting operator is the middle composition \composegap{m}{g}{m}. 
%
When the first middle is a failure (\hcmfail{l}), the result is the second
middle. In all the remaining cases, the first middle is not a failure.
%
When the gap is a cast between shallowly inconsistent types (i.e. having 
different top-most constructors). the output is a failure.
%
When the gap is good but the second middle is a failure, the output is the same 
failure. A gap $g$ is good if and only if there is not $P$ and $l$ 
and $Q$ such that $g = \gsome{P}{l}{Q}$ and \notshallowlyconsistent{P}{Q}.
%
The remaining cases are straightforward recursion.

Agda is able to understand that \composegap{c}{g}{c} is terminating. The reason 
might be that in \composegap{c}{g}{c} both casts become smaller in recursive 
calls, although their position is swapped in the function case. Readers who
hate non-structural recursion might add a Boolean parameter to indicate the order
of composition.

\begin{figure}
	\fbox{$\translate{\ccast{S}{l}{T}} = c$} Translation of \LD\ hypercoercions
	\[
	\begin{array}{rclr}
	\figequalto{\translate{\ccast{S}{l}{T}}}{\composegap{id(S)}{(\gsome{S}{l}{T})}{id(T)}}{}
	\end{array}
	\]
	\caption{Translatin of \LD\ Hypercoercions}
	\label{fig:LazyD-Hypercoercions-translation}
\end{figure}

The definition of translation \translate{\ccast{S}{l}{T}} is straightforward 
(\figref{fig:LazyD-Hypercoercions-translation}).

\begin{figure}
\fbox{$\denote{c}(v)=r$}
\[
\begin{array}{rclr}
\figequalto{\denote{\hcci}(v)}{v}{}
\figequalto{\denote{\hccc{h}{m}{t}}(v)}{
	(\denote{h}_h(v) \mbind \denote{m}_m)
	\mbind \denote{t}_t
}{}
\end{array}
\]

\fbox{$\denote{h}_h(v)=r$}
\[
\begin{array}{rclr}
\figequalto{\denote{\hche}(v)}{v}{}
\figequalto{\denote{\hchp{Q}{l}}(\vdyn{P}{u})}{
	\denote{\composegap{\idm{P}}{\gsome{P}{l}{Q}}{\idm{Q}}}_m(u)
}{}
\end{array}
\]

\fbox{$\denote{m}_m(v)=r$}
\[
\begin{array}{rclr}
\figequalto{\denote{\hcmfail{l}}(v)}{\error{l}}{}
\figequalto{\denote{\hcmbool}(v)}{v}{}
\figequalto{\denote{\hcmfunc{d_1}{d_2}}(\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{c_1}{c_2}})}{
	\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{(\compose{d_1}{c_1})}{(\compose{c_2}{d_2})}}
}{}
\figequalto{\denote{\hcmprod{d_1}{d_2}}(\ecast{\vcons{v_1}{v_2}}{\Pprod{c_1}{c_2}})}{
	\ecast{\vcons{v_1}{v_2}}{\Pprod{(\compose{c_1}{d_1})}{(\compose{c_2}{d_2})}}
}{}
\end{array}
\]

\fbox{$\denote{t}_t(v)=r$}
\[
\begin{array}{rclr}
\figequalto{\denote{\hcte}(v)}{v}{}
\figequalto{\denote{\hcti{P}}(v)}{\vdyn{P}{v}}{}
\end{array}
\]
\caption{Application of \LD\ Hypercoercions}
\label{fig:LazyD-Hypercoercions-application}
\end{figure}

Fig.~\ref{fig:LazyD-Hypercoercions-application} defines the cast application 
function for \LD\ hypercoercions.
Applying the identity cast for the dynamic type succeeds immediately. 
Otherwise, we apply head, middle, and tail in sequence. 
%
In projection case of head application, a new middle is constructed and applied 
to the injected value.
%
Middle applications raises an error if the middle is a failure and extend value 
proxies otherwise.

\begin{proposition} Lazy D Hypercoercions implement the Lazy D Cast ADT
\end{proposition}
\begin{theorem}[Lazy D Hypercoercions are Correct] Let $H$ be the Lazy D 
hypercoercions,
\[\machineequiv{\LDMachine}{\SMachine{H}}\]
\end{theorem}

\subsection{\LUD\ Hypercoercions} \label{sec:LazyUD-Hypercoercions}

Hypercoercions for the \LUD\ blame tracking strategy have the same syntax
as for \LD\ (\figref{fig:hypercoercion}), but the definitions of $\id{T}$ and
$\translate{\ccast{S}{l}{T}}$ and $\compose{c}{c}$, differ from \LD. Here 
composition $\compose{c_1}{c_2}$ makes the usual assumption that the
target type of $c_1$ matches the source type of $c_2$. 
%
The definition of composition for \LUD\ is particularly straightforward.
The first two lines say that \hcci\ acts as the identity on both the left
and right.
The third line handles the case when both hypercoercions have three parts. 
%
The first three lines of the middle composition $m \fatsemi t \fatsemi h 
\fatsemi m$ function match its arguments in sequential order to determine which 
label to blame. The forth and last line is straightforward recursion.

\todo[inline]{KC: I need a better name for \compatibletailhead{t}{h}}

%The third line handles the case when both the tail of $c_1$
%and the head of $c_2$ are no-ops, in which case the middles types
%$m_1$ and $m_2$ are composed via an auxiliary composition
%operator. This compose operator for middle coercions assumes that its
%inputs are shallowly consistent.  The fourth line handles the
%important case when the tail of $c_1$ is an injection and the head of
%$c_2$ is a projection. If the two middle coercions are shallowly
%consistent, then they can be composed. If not, the tail of the result
%is a failure coercion with the projection's label.  The last line
%handles the case when the tail of $c_1$ is a failure, in which case
%the result of composition is $c_1$. The definition of $id$ is straightforward.

\begin{figure}
	\fbox{$c \fatsemi c = c$}
	\[
	\begin{array}{rclcl}
		\hcci          
		& \fatsemi & d
		& = & d \\
		\hccc{h}{m}{t}
		& \fatsemi & \hcci
		& = & \hccc{h}{m}{t} \\
		\hccc{h_1}{m_1}{t_1}
		& \fatsemi & \hccc{h_2}{m_2}{t_2}
		& = & \hccc{h_1}{(m_1 \fatsemi t_1 \fatsemi h_2 \fatsemi m_2)}{t_2}
	\end{array}
	\]

	\fbox{$m \fatsemi t \fatsemi h \fatsemi m = m$}
	\[
	\begin{array}{rccccclclr}
	\hcmfail{l} 
	& \fatsemi & t
	& \fatsemi & h
	& \fatsemi & m_2
	& = & \hcmfail{l} \\
	m_1
	& \fatsemi & \hcti{G}
	& \fatsemi & \hchp{H}{l}
	& \fatsemi & m_2
	& = & \hcmfail{l}
	& \sidecond{G \neq H \text{ and } \forall l. m_1 \neq \hcmfail{l}} \\
	m_1
	& \fatsemi & t
	& \fatsemi & h
	& \fatsemi & \hcmfail{l}
	& = & \hcmfail{l}
	& \sidecond{\compatibletailhead{t}{h}
		\text{ and } \forall l. m_1 \neq \hcmfail{l}} \\
	m_1
	& \fatsemi & t
	& \fatsemi & h
	& \fatsemi & m_2
	& = & m_1 \fatsemi m_2
	& \sidecond{
		\forall l. m_2 \neq \hcmfail{l}
		\text{ and } \compatibletailhead{t}{h}
		\text{ and } \forall l. m_1 \neq \hcmfail{l}} \\
	\end{array}
	\]

	\fbox{$m \fatsemi m = m$}
	\[
	\begin{array}{rclcl}
	\hcmbool
	& \fatsemi & \hcmbool
	& = & \hcmbool \\
	\hcmfunc{c_1}{d_1}
	& \fatsemi & \hcmfunc{c_2}{d_2}
	& = & \hcmfunc{c_2 \fatsemi c_1}{d_1 \fatsemi d_2} \\
	\hcmprod{c_1}{d_1}
	& \fatsemi & \hcmprod{c_2}{d_2}
	& = & \hcmprod{c_2 \fatsemi c_1}{d_1 \fatsemi d_2} \\
	\end{array}
	\]

	\fbox{\compatibletailhead{t}{h}}
	\begin{gather*}
	\infr{}{\compatibletailhead{\hcte}{\hche}}
	\quad
	\infr{}{\compatibletailhead{\hcti{G}}{\hchp{G}{l}}}
	\end{gather*}

\caption{\LUD\ Hypercoercions (1): composition}
\label{fig:LazyUD-Hypercoercions-composition}
\end{figure}

\begin{figure}
	\fbox{$\translate{\ccast{S}{l}{T}} = c$}
	\[
	\begin{array}{rcll}
	\figequalto{\translate{\ccast{\Tdyn}{l}{\Tdyn}}}{\hcci}{}
	\figequalto{\translate{\ccast{\Tdyn}{l}{Q}}}{\castFromDyn{l}{Q}}{}
	\figequalto{\translate{\ccast{P}{l}{\Tdyn}}}{\castToDyn{P}{l}}{}
	\figequalto{\translate{\ccast{P}{l}{Q}}}{
		\hccc{\hche}{\hcmfail{l}}{\hcte}
	}{\sidecond{\notshallowlyconsistent{P}{Q}}}
	\figequalto{\translate{\ccast{\Pbool}{l}{\Pbool}}}{
		\hccc{\hche}{\hcmbool}{\hcte}
	}{}
	\figequalto{\translate{\ccast{\Pfunc{S_1}{T_1}}{l}{\Pfunc{S_2}{T_2}}}}{
		\hccc{\hche}{
			\hcmfunc{
				\translate{\ccast{S_2}{l}{S_1}}}{
				\translate{\ccast{T_1}{l}{T_2}}}
		}{\hcte}
	}{}
	\figequalto{\translate{\ccast{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}{
		\hccc{\hche}{
			\hcmprod{
				\translate{\ccast{S_1}{l}{S_2}}}{
				\translate{\ccast{T_1}{l}{T_2}}}
		}{\hcte}
	}{}
	\end{array}
	\]

	\fbox{$\castFromDyn{Q}{l}$}
	\[
	\begin{array}{rcll}
\figequalto{\castFromDyn{\Tdyn}{l}}{\hcci}{}
\figequalto{\castFromDyn{\Pbool}{l}}{
	\hccc{\hchp{\Gbool}{l}}{
		\hcmbool
	}{\hcte}
}{}
\figequalto{\castFromDyn{\Pfunc{S}{T}}{l}}{
	\hccc{\hchp{\Gfunc}{l}}{
		\hcmfunc{\castToDyn{S}{l}}{\castFromDyn{T}{l}}
	}{\hcte}
}{}
\figequalto{\castFromDyn{\Pprod{S}{T}}{l}}{
	\hccc{\hchp{\Gprod}{l}}{
		\hcmprod{\castFromDyn{S}{l}}{\castFromDyn{T}{l}}
	}{\hcte}
}{}
	\end{array}
	\]

	\fbox{$\castToDyn{Q}{l}$}
	\[
	\begin{array}{rcll}
\figequalto{\castToDyn{\Tdyn}{l}}{\hcci}{}
\figequalto{\castToDyn{\Pbool}{l}}{
	\hccc{\hche}{
		\hcmbool
	}{\hcti{\Gbool}}
}{}
\figequalto{\castToDyn{\Pfunc{S}{T}}{l}}{
	\hccc{\hche}{
		\hcmfunc{\castFromDyn{S}{l}}{\castToDyn{T}{l}}
	}{\hcti{\Gfunc}}
}{}
\figequalto{\castToDyn{\Pprod{S}{T}}{l}}{
	\hccc{\hche}{
		\hcmprod{\castToDyn{S}{l}}{\castToDyn{T}{l}}
	}{\hcti{\Gprod}}
}{}
	\end{array}
	\]
\caption{\LUD\ Hypercoercions (2): translation}
\label{fig:LazyUD-Hypercoercions-translation}
\end{figure}

The translation function \translate{\ccast{S}{l}{T}} is defined in 
\figref{fig:LazyUD-Hypercoercions-translation}.  Prior
presentations of this function do not use auxiliary functions, as we
do here. The reason that we introduce the auxiliary functions
$\mathit{castToDyn}$ and $\mathit{castFromDyn}$, is to ensure that each of them 
is structurally recursive, which makes them straightforward to define in
Agda.

\begin{figure}
	\fbox{$\denote{c}_c(v)=r$}
	\[
	\begin{array}{rclr}
	\figequalto{\denote{\hcci}(v)}{v}{}
	\figequalto{\denote{\hccc{h}{m}{t}}(v)}{
		(\denote{h}(v) \mbind \denote{m})
		\mbind \denote{t}
	}{}
	\end{array}
	\]

	\fbox{$\denote{h}_h(v)=r$}
	\[
	\begin{array}{rclr}
	\figequalto{\denote{\hche}(v)}{v}{}
	\figequalto{\denote{\hchp{G}{l}}(\vdyn{G}{u})}{u}{}
	\figequalto{\denote{\hchp{H}{l}}(\vdyn{G}{u})}{\error{l}}{
		\sidecond{G \neq H}
	}
	\end{array}
	\]

	\fbox{$\denote{m}_m(v)=r$}
	\[
	\begin{array}{rclr}
	\figequalto{\denote{\hcmfail{l}}(v)}{\error{l}}{}
	\figequalto{\denote{\hcmbool}(v)}{v}{}
	\figequalto{\denote{\hcmfunc{d_1}{d_2}}(\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{c_1}{c_2}})}{
		\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{(\compose{d_1}{c_1})}{(\compose{c_2}{d_2})}}
	}{}
	\figequalto{\denote{\hcmprod{d_1}{d_2}}(\ecast{\vcons{v_1}{v_2}}{\Pprod{c_1}{c_2}})}{
		\ecast{\vcons{v_1}{v_2}}{\Pprod{(\compose{c_1}{d_1})}{(\compose{c_2}{d_2})}}
	}{}
	\end{array}
	\]

	\fbox{$\denote{t}_t(v)=r$}
	\[
	\begin{array}{rclr}
	\figequalto{\denote{\hcte}(v)}{v}{}
	\figequalto{\denote{\hcti{G}}(v)}{\vdyn{G}{v}}{}
	\end{array}
	\]
	\caption{Application of \LUD\ Hypercoercions}
	\label{fig:LazyUD-Hypercoercions-application}
\end{figure}

Fig.~\ref{fig:LazyUD-Hypercoercions-application} defines the cast application 
function for \LUD\ hypercoercions.
Applying the identity cast for the dynamic type succeeds immediately. 
Otherwise, we apply head, middle, and tail in sequence. 
%
In the two projection cases of head application, the result depends on whether 
the source ground type ($G$) is the same as the target ($H$).
%
Middle applications raises an error if the middle is a failure and extend value 
proxies otherwise.

\begin{proposition} Lazy UD Hypercoercions implement the Lazy UD Cast ADT
\end{proposition}
\begin{theorem}[Lazy UD Hypercoercions are Correct] Let $H$ be the Lazy UD 
	hypercoercions,
	\[\machineequiv{\LDMachine}{\SMachine{H}}\]
\end{theorem}

\subsection{Compact Representation of Hypercoercions} \label{sec:hypercoercions-compact}

Hypercoercions enable a bit-level representation that is particularly
compact for identity coercions and coercions that project from base
types or inject to base types.  (Here \Pbool\ is the only base
type, but in a real language the base types would include integers,
Booleans, etc.)  We conjecture that such coercions occur more
frequently than the more complex coercions (e.g. between function
types), especially because such coercions appear in the leaves of
complex coercions.

Furthermore, values of base types are often stored in CPU registers, so
it would be nice for coercions on base types to also fit in registers,
that is, in 64 bits, so that applying a coercion to a value of base
type would not require access to main memory, which is an order of
magnitude slower than accessing registers on a modern CPU.
%
Coercions involving non-base types, such as function types, may be
arbitrarily deep, so in those cases, the hypercoercion representation
has to be a pointer to a heap-allocated structure.

Here is a sketch for the bit-level representation for hypercoercions.

\begin{itemize}
\item 1 bit to differentiate between \hcci\ and three-part coercions.

\item 1 bit to differentiate between middle coercions of base type
versus non-base type.

\item If the middle is a non-base type or bottom ($\hcmfail{l}$), then 61 bits 
represent a
pointer to a heap-allocated structure. Heap allocated structures are
usually 8-byte aligned, so there are 3 unused bits in a pointer.

\item If the middle is of base types, then the remaining 62 bits are
used to represent the head (51 bits), middle (10 bit), and tail (1
bit).
\begin{itemize}
\item The head requires 1 bit to differentiate between $\hche$
and $\hchp{P}{l}$ and then 50 bits could be used for the label $l$.
\item The middle would use 10 bits to differentiate all the base types.
\item The tail requires 1 bits to differentiate between $\hcte$ and $\hcti{P}$, 
where $P$ must be equal to the middle type.
\end{itemize}
The bits for blame labels represent an index into a table of blame information.
In the event that a program requires too many blame labels, then the 
implementation
can fallback to using the heap-allocated structure for more
coercions.
\end{itemize}

\section{Conclusion} \label{sec:conclusion}

We present steps toward the first framework for proving the
correctness of cast representations.
%
The framework includes abstract data types for cast representations
(Cast ADT and its refinements Lazy D Cast ADT and \LUD\ \CastADT), 
a parameterized abstract machine \effCEK{C}, 
a theorem which states that \SMachine{C} is equivalent
to \LDMachine\ when $C$ is a Lazy D Cast ADT, and
a theorem which states that \SMachine{C} is equivalent
to \LUDMachine\ when $C$ is a Lazy UD Cast ADT.
%
%The Lazy D Cast ADT refines the Cast ADT. 
%
We applied this framework to Lazy UD coercions in normal form,
Lazy UD threesomes, and hypercoercions and conjecture that
this framework is also applicable to other cast representations.

We also presented hypercoercions, a cast representation that has a 
theorem-prover-friendly compose operator and a more compact memory
representation in comparison with coercions in normal forms.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\bibliographystyle{splncs04}
\bibliography{bibfile,all-modified}{}

\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}